---
title: "Covid plots for family"
output:
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 3
  pdf_document:
    toc: true
    toc_depth: 3
date: '`r format(Sys.time(), "%B %d, %Y")`'
---

## Background
* This page shows COVID data broken down into regions that affect our family. It pulls from publicly available raw data from the NY Times, which is reported daily. It's on a one-day lag, so when you see this report, the most recent day included while be the day prior.
* The page is mostly organized by geography: It starts by showing data for all of the U.S., followed by our respective states, followed by our respective counties.
* Within each geophraphy, it shows several plots:
  * Total confirmed cases: This is the cumulative total of cases reported.
  * New confirmed cases: This is the number of new cases reported on each respective date.
  * Rolling 3-day case growth rate: This is the average daily growth rate in total cases over the previous 3 days. For instance, if it is 100%, then that means total cases are doubling every 3 days. (This isn't shown at the county level because the sample size is too small.)
  * Total deaths: This is the cumulative total of deaths reported
  * New deaths: This is the number of new deaths reported on each respective date
  * Rolling 3-day death growth rate: This is the average daily growth rate in total deaths over the previous 3 days. For instance, if it is 100%, then that means total deaths are doubling every 3 days. (This isn't shown at the county level because the sample size is too small.)
* Besides this data for each geography, it also includes some additional plots at the end:
  * It shows tables of both total cases and deaths *per million residents* for all states, and then a graph of trends of these numbers for the top 10 states (only using 10 in the graph so it's not too crowded). Showing the numbers per million helps normalize for the size of the states when looking at totals.
  * It then shows these same tables and plots for cases and deaths, but with new cases/deaths per million residents instead of total cases/deaths per million. This section uses a rolling 3-day average of new cases/deaths.
  * Finally, it shows the same rolling 3-day average of new cases per million residents, but broken out by states that have enacted stay-at-home orders and when the orders were enacted.
  

```{r include = F, results = 'hide'}
# load packages
library(data.table)
library(ggplot2)
library(scales)
library(dplyr)
library(lubridate)
library(utils)
library(httr)
library(graphics)

# turn off scientific notation
options(scipen = 999)
```

```{r include = F, results = 'hide'}
state.raw <- data.table(read.csv(file=url("https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-states.csv"), header = T))
county.raw <- data.table(read.csv(file=url("https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-counties.csv"), header = T))
```

```{r include = F, results = 'hide'}
# format dates
state.raw[, date := as.Date(date)]
county.raw[, date := as.Date(date)]

# grad stay-at-home dates (manually updated locally)
stay_dates <- fread("~/covid19/stay_at_home_dates.csv", header = T)
stay_dates[, start_date := as.Date(start_date, format = "%Y-%m-%d")]
stay_dates[, population := as.integer(gsub(",", "", population))]
```

## United States

### Confirmed cases

```{r echo = F}
agg.cases.US <- state.raw[order(date),
                          .(total_cases = sum(cases, na.rm = T)),
                          by = "date"]
agg.cases.US.plot <- ggplot(agg.cases.US[date >= "2020-03-01"], aes(x = date, y = total_cases)) + 
  geom_bar(stat = "identity") +
    labs(x = "Date",
       y = "Total confirmed COVID cases",
       title = "Total confirmed COVID cases",
       subtitle = "United States") +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) +
  scale_y_continuous(labels = comma)
agg.cases.US.plot
```

```{r echo = F}
agg.cases.US <- state.raw[order(date),
                          .(total_cases = sum(cases, na.rm = T)),
                          by = "date"]
temp <- agg.cases.US
agg.cases.US[, total_cases_prev := temp$total_cases[match(agg.cases.US$date - 1, temp$date)]]
agg.cases.US[, total_new_cases := total_cases - total_cases_prev]
agg.new.cases.US.plot <- ggplot(agg.cases.US[date >= "2020-03-01"], aes(x = date, y = total_new_cases)) + 
  geom_bar(stat = "identity") +
    labs(x = "Date",
       y = "Confirmed new COVID cases",
       title = "Confirmed new COVID cases by day",
       subtitle = "United States") +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) +
  scale_y_continuous(labels = comma)
agg.new.cases.US.plot
```

```{r echo = F}
agg.cases.US <- state.raw[order(date),
                          .(total_cases = sum(cases, na.rm = T)),
                          by = "date"]
temp <- agg.cases.US
agg.cases.US[, total_cases_prev3 := temp$total_cases[match(agg.cases.US$date - 3, temp$date)]]
agg.cases.US[, total_cases_rtgrowth := total_cases/total_cases_prev3 - 1]
agg.cases.growth.US.plot <- ggplot(agg.cases.US[date >= "2020-03-01"], aes(x = date, y = total_cases_rtgrowth)) + 
  geom_line() +
    labs(x = "Date",
       y = "Rate of growth (rolling 3-day window)",
       title = "Rolling 3-day rate of growth: Total confirmed COVID cases",
       subtitle = "United States") +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1), limits = c(0, max(agg.cases.US$total_cases_rtgrowth * 1.1)))
agg.cases.growth.US.plot
```

### Deaths

```{r echo = F}
agg.deaths.US <- state.raw[order(date),
                          .(total_deaths = sum(deaths, na.rm = T)),
                          by = "date"]
agg.deaths.US.plot <- ggplot(agg.deaths.US[date >= "2020-03-01"], aes(x = date, y = total_deaths)) + 
  geom_bar(stat = "identity") +
    labs(x = "Date",
       y = "Total COVID-related deaths",
       title = "Total COVID-related deaths",
       subtitle = "United States") +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) +
  scale_y_continuous(labels = comma)
agg.deaths.US.plot
```

```{r echo = F}
agg.deaths.US <- state.raw[order(date),
                          .(total_deaths = sum(deaths, na.rm = T)),
                          by = "date"]
temp <- agg.deaths.US
agg.deaths.US[, total_deaths_prev := temp$total_deaths[match(agg.deaths.US$date - 1, temp$date)]]
agg.deaths.US[, total_new_deaths := total_deaths - total_deaths_prev]
agg.new.deaths.US.plot <- ggplot(agg.deaths.US[date >= "2020-03-01"], aes(x = date, y = total_new_deaths)) + 
  geom_bar(stat = "identity") +
    labs(x = "Date",
       y = "New COVID-related deaths",
       title = "New COVID-related deaths by day",
       subtitle = "United States") +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) +
  scale_y_continuous(labels = comma)
agg.new.deaths.US.plot
```

```{r echo = F}
agg.deaths.US <- state.raw[order(date),
                          .(total_deaths = sum(deaths, na.rm = T)),
                          by = "date"]
temp <- agg.deaths.US
agg.deaths.US[, total_deaths_prev3 := temp$total_deaths[match(agg.deaths.US$date - 3, temp$date)]]
agg.deaths.US[, total_deaths_rtgrowth := total_deaths/total_deaths_prev3 - 1]
agg.deaths.growth.US.plot <- ggplot(agg.deaths.US[date >= "2020-03-10"], aes(x = date, y = total_deaths_rtgrowth)) + 
  geom_line() +
    labs(x = "Date",
       y = "Rate of growth (rolling 3-day window)",
       title = "Rolling 3-day rate of growth: Total COVID-related deaths",
       subtitle = "United States") +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1), limits = c(0, max(agg.deaths.US$total_deaths_rtgrowth * 1.1)))
agg.deaths.growth.US.plot
```

## Washington

### Confirmed cases

```{r echo = F}
# adjust state name to flow through rest of code (use capitalized full state name)
use.state <- "Washington"
cases.state <- state.raw[state %in% use.state]
agg.cases.state <- cases.state[order(date),
                          .(total_cases = sum(cases, na.rm = T)),
                          by = "date"]
agg.cases.state.plot <- ggplot(agg.cases.state[date >= "2020-03-01"], aes(x = date, y = total_cases)) + 
  geom_bar(stat = "identity") +
    labs(x = "Date",
       y = "Total confirmed COVID cases",
       title = "Total confirmed COVID cases",
       subtitle = use.state) +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) +
  scale_y_continuous(labels = comma)
agg.cases.state.plot
```

```{r echo = F}
use.state <- "Washington"
cases.state <- state.raw[state %in% use.state]
agg.cases.state <- cases.state[order(date),
                          .(total_cases = sum(cases, na.rm = T)),
                          by = "date"]
temp <- agg.cases.state
agg.cases.state[, total_cases_prev := temp$total_cases[match(agg.cases.state$date - 1, temp$date)]]
agg.cases.state[, total_new_cases := total_cases - total_cases_prev]
agg.new.cases.state.plot <- ggplot(agg.cases.state[date >= "2020-03-01"], aes(x = date, y = total_new_cases)) + 
  geom_bar(stat = "identity") +
    labs(x = "Date",
       y = "Confirmed new COVID cases",
       title = "Confirmed new COVID cases by day",
       subtitle = use.state) +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) +
  scale_y_continuous(labels = comma)
agg.new.cases.state.plot
```

```{r echo = F}
use.state <- "Washington"
cases.state <- state.raw[state %in% use.state]
agg.cases.state <- cases.state[order(date),
                          .(total_cases = sum(cases, na.rm = T)),
                          by = "date"]
temp <- agg.cases.state
agg.cases.state[, total_cases_prev3 := temp$total_cases[match(agg.cases.state$date - 3, temp$date)]]
agg.cases.state[, total_cases_rtgrowth := total_cases/total_cases_prev3 - 1]
agg.cases.growth.state.plot <- ggplot(agg.cases.state[date >= "2020-03-01"], aes(x = date, y = total_cases_rtgrowth)) + 
  geom_line() +
    labs(x = "Date",
       y = "Rate of growth (rolling 3-day window)",
       title = "Rolling 3-day rate of growth: Total confirmed COVID cases",
       subtitle = use.state) +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1), limits = c(0, max(agg.cases.state$total_cases_rtgrowth * 1.1)))
agg.cases.growth.state.plot
```

### Deaths

```{r echo = F}
# adjust state name to flow through rest of code (use capitalized full state name)
use.state <- "Washington"
deaths.state <- state.raw[state %in% use.state]
agg.deaths.state <- cases.state[order(date),
                          .(total_deaths = sum(deaths, na.rm = T)),
                          by = "date"]
agg.deaths.state.plot <- ggplot(agg.deaths.state[date >= "2020-03-01"], aes(x = date, y = total_deaths)) + 
  geom_bar(stat = "identity") +
    labs(x = "Date",
       y = "Total COVID-related deaths",
       title = "Total COVID-related deaths",
       subtitle = use.state) +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))
agg.deaths.state.plot
```

```{r echo = F}
use.state <- "Washington"
deaths.state <- state.raw[state %in% use.state]
agg.deaths.state <- deaths.state[order(date),
                          .(total_deaths = sum(deaths, na.rm = T)),
                          by = "date"]
temp <- agg.deaths.state
agg.deaths.state[, total_deaths_prev := temp$total_deaths[match(agg.deaths.state$date - 1, temp$date)]]
agg.deaths.state[, total_new_deaths := total_deaths - total_deaths_prev]
agg.new.deaths.state.plot <- ggplot(agg.deaths.state[date >= "2020-03-01"], aes(x = date, y = total_new_deaths)) + 
  geom_bar(stat = "identity") +
    labs(x = "Date",
       y = "New COVID-related deaths",
       title = "New COVID-related deaths by day",
       subtitle = use.state) +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))
agg.new.deaths.state.plot
```

```{r echo = F}
use.state <- "Washington"
deaths.state <- state.raw[state %in% use.state]
agg.deaths.state <- deaths.state[order(date),
                          .(total_deaths = sum(deaths, na.rm = T)),
                          by = "date"]
temp <- agg.deaths.state
agg.deaths.state[, total_deaths_prev3 := temp$total_deaths[match(agg.deaths.state$date - 3, temp$date)]]
agg.deaths.state[, total_deaths_rtgrowth := total_deaths/total_deaths_prev3 - 1]
agg.deaths.growth.state.plot <- ggplot(agg.deaths.state[date >= "2020-03-10"], aes(x = date, y = total_deaths_rtgrowth)) + 
  geom_line() +
    labs(x = "Date",
       y = "Rate of growth (rolling 3-day window)",
       title = "Rolling 3-day rate of growth: Total COVID-related deaths",
       subtitle = use.state) +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1), limits = c(0, max(agg.deaths.state$total_deaths_rtgrowth * 1.1)))
agg.deaths.growth.state.plot
```

## California

### Confirmed cases

```{r echo = F}
# adjust state name to flow through rest of code (use capitalized full state name)
use.state <- "California"
cases.state <- state.raw[state %in% use.state]
agg.cases.state <- cases.state[order(date),
                          .(total_cases = sum(cases, na.rm = T)),
                          by = "date"]
agg.cases.state.plot <- ggplot(agg.cases.state[date >= "2020-03-01"], aes(x = date, y = total_cases)) + 
  geom_bar(stat = "identity") +
    labs(x = "Date",
       y = "Total confirmed COVID cases",
       title = "Total confirmed COVID cases",
       subtitle = use.state) +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) +
  scale_y_continuous(labels = comma)
agg.cases.state.plot
```

```{r echo = F}
use.state <- "California"
cases.state <- state.raw[state %in% use.state]
agg.cases.state <- cases.state[order(date),
                          .(total_cases = sum(cases, na.rm = T)),
                          by = "date"]
temp <- agg.cases.state
agg.cases.state[, total_cases_prev := temp$total_cases[match(agg.cases.state$date - 1, temp$date)]]
agg.cases.state[, total_new_cases := total_cases - total_cases_prev]
agg.new.cases.state.plot <- ggplot(agg.cases.state[date >= "2020-03-01"], aes(x = date, y = total_new_cases)) + 
  geom_bar(stat = "identity") +
    labs(x = "Date",
       y = "Confirmed new COVID cases",
       title = "Confirmed new COVID cases by day",
       subtitle = use.state) +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) +
  scale_y_continuous(labels = comma)
agg.new.cases.state.plot
```

```{r echo = F}
use.state <- "California"
cases.state <- state.raw[state %in% use.state]
agg.cases.state <- cases.state[order(date),
                          .(total_cases = sum(cases, na.rm = T)),
                          by = "date"]
temp <- agg.cases.state
agg.cases.state[, total_cases_prev3 := temp$total_cases[match(agg.cases.state$date - 3, temp$date)]]
agg.cases.state[, total_cases_rtgrowth := total_cases/total_cases_prev3 - 1]
agg.cases.growth.state.plot <- ggplot(agg.cases.state[date >= "2020-03-01"], aes(x = date, y = total_cases_rtgrowth)) + 
  geom_line() +
    labs(x = "Date",
       y = "Rate of growth (rolling 3-day window)",
       title = "Rolling 3-day rate of growth: Total confirmed COVID cases",
       subtitle = use.state) +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1), limits = c(0, max(agg.cases.state$total_cases_rtgrowth * 1.1)))
agg.cases.growth.state.plot
```

### Deaths

```{r echo = F}
# adjust state name to flow through rest of code (use capitalized full state name)
use.state <- "California"
deaths.state <- state.raw[state %in% use.state]
agg.deaths.state <- cases.state[order(date),
                          .(total_deaths = sum(deaths, na.rm = T)),
                          by = "date"]
agg.deaths.state.plot <- ggplot(agg.deaths.state[date >= "2020-03-01"], aes(x = date, y = total_deaths)) + 
  geom_bar(stat = "identity") +
    labs(x = "Date",
       y = "Total COVID-related deaths",
       title = "Total COVID-related deaths",
       subtitle = use.state) +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))
agg.deaths.state.plot
```

```{r echo = F}
use.state <- "California"
deaths.state <- state.raw[state %in% use.state]
agg.deaths.state <- deaths.state[order(date),
                          .(total_deaths = sum(deaths, na.rm = T)),
                          by = "date"]
temp <- agg.deaths.state
agg.deaths.state[, total_deaths_prev := temp$total_deaths[match(agg.deaths.state$date - 1, temp$date)]]
agg.deaths.state[, total_new_deaths := total_deaths - total_deaths_prev]
agg.new.deaths.state.plot <- ggplot(agg.deaths.state[date >= "2020-03-01"], aes(x = date, y = total_new_deaths)) + 
  geom_bar(stat = "identity") +
    labs(x = "Date",
       y = "New COVID-related deaths",
       title = "New COVID-related deaths by day",
       subtitle = use.state) +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))
agg.new.deaths.state.plot
```

```{r echo = F}
use.state <- "California"
deaths.state <- state.raw[state %in% use.state]
agg.deaths.state <- deaths.state[order(date),
                          .(total_deaths = sum(deaths, na.rm = T)),
                          by = "date"]
temp <- agg.deaths.state
agg.deaths.state[, total_deaths_prev3 := temp$total_deaths[match(agg.deaths.state$date - 3, temp$date)]]
agg.deaths.state[, total_deaths_rtgrowth := total_deaths/total_deaths_prev3 - 1]
agg.deaths.growth.state.plot <- ggplot(agg.deaths.state[date >= "2020-03-10"], aes(x = date, y = total_deaths_rtgrowth)) + 
  geom_line() +
    labs(x = "Date",
       y = "Rate of growth (rolling 3-day window)",
       title = "Rolling 3-day rate of growth: Total COVID-related deaths",
       subtitle = use.state) +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1), limits = c(0, max(agg.deaths.state$total_deaths_rtgrowth * 1.1)))
agg.deaths.growth.state.plot
```

## South Carolina

### Confirmed cases

```{r echo = F}
# adjust state name to flow through rest of code (use capitalized full state name)
use.state <- "South Carolina"
cases.state <- state.raw[state %in% use.state]
agg.cases.state <- cases.state[order(date),
                          .(total_cases = sum(cases, na.rm = T)),
                          by = "date"]
agg.cases.state.plot <- ggplot(agg.cases.state[date >= "2020-03-01"], aes(x = date, y = total_cases)) + 
  geom_bar(stat = "identity") +
    labs(x = "Date",
       y = "Total confirmed COVID cases",
       title = "Total confirmed COVID cases",
       subtitle = use.state) +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) +
  scale_y_continuous(labels = comma)
agg.cases.state.plot
```

```{r echo = F}
use.state <- "South Carolina"
cases.state <- state.raw[state %in% use.state]
agg.cases.state <- cases.state[order(date),
                          .(total_cases = sum(cases, na.rm = T)),
                          by = "date"]
temp <- agg.cases.state
agg.cases.state[, total_cases_prev := temp$total_cases[match(agg.cases.state$date - 1, temp$date)]]
agg.cases.state[, total_new_cases := total_cases - total_cases_prev]
agg.new.cases.state.plot <- ggplot(agg.cases.state[date >= "2020-03-01"], aes(x = date, y = total_new_cases)) + 
  geom_bar(stat = "identity") +
    labs(x = "Date",
       y = "Confirmed new COVID cases",
       title = "Confirmed new COVID cases by day",
       subtitle = use.state) +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) +
  scale_y_continuous(labels = comma)
agg.new.cases.state.plot
```

```{r echo = F}
use.state <- "South Carolina"
cases.state <- state.raw[state %in% use.state]
agg.cases.state <- cases.state[order(date),
                          .(total_cases = sum(cases, na.rm = T)),
                          by = "date"]
temp <- agg.cases.state
agg.cases.state[, total_cases_prev3 := temp$total_cases[match(agg.cases.state$date - 3, temp$date)]]
agg.cases.state[, total_cases_rtgrowth := total_cases/total_cases_prev3 - 1]
agg.cases.growth.state.plot <- ggplot(agg.cases.state[date >= "2020-03-01"], aes(x = date, y = total_cases_rtgrowth)) + 
  geom_line() +
    labs(x = "Date",
       y = "Rate of growth (rolling 3-day window)",
       title = "Rolling 3-day rate of growth: Total confirmed COVID cases",
       subtitle = use.state) +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1), limits = c(0, max(agg.cases.state$total_cases_rtgrowth * 1.1)))
agg.cases.growth.state.plot
```

### Deaths

```{r echo = F}
# adjust state name to flow through rest of code (use capitalized full state name)
use.state <- "South Carolina"
deaths.state <- state.raw[state %in% use.state]
agg.deaths.state <- cases.state[order(date),
                          .(total_deaths = sum(deaths, na.rm = T)),
                          by = "date"]
agg.deaths.state.plot <- ggplot(agg.deaths.state[date >= "2020-03-01"], aes(x = date, y = total_deaths)) + 
  geom_bar(stat = "identity") +
    labs(x = "Date",
       y = "Total COVID-related deaths",
       title = "Total COVID-related deaths",
       subtitle = use.state) +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))
agg.deaths.state.plot
```

```{r echo = F}
use.state <- "South Carolina"
deaths.state <- state.raw[state %in% use.state]
agg.deaths.state <- deaths.state[order(date),
                          .(total_deaths = sum(deaths, na.rm = T)),
                          by = "date"]
temp <- agg.deaths.state
agg.deaths.state[, total_deaths_prev := temp$total_deaths[match(agg.deaths.state$date - 1, temp$date)]]
agg.deaths.state[, total_new_deaths := total_deaths - total_deaths_prev]
agg.new.deaths.state.plot <- ggplot(agg.deaths.state[date >= "2020-03-01"], aes(x = date, y = total_new_deaths)) + 
  geom_bar(stat = "identity") +
    labs(x = "Date",
       y = "New COVID-related deaths",
       title = "New COVID-related deaths by day",
       subtitle = use.state) +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))
agg.new.deaths.state.plot
```

```{r echo = F}
use.state <- "South Carolina"
deaths.state <- state.raw[state %in% use.state]
agg.deaths.state <- deaths.state[order(date),
                          .(total_deaths = sum(deaths, na.rm = T)),
                          by = "date"]
temp <- agg.deaths.state
agg.deaths.state[, total_deaths_prev3 := temp$total_deaths[match(agg.deaths.state$date - 3, temp$date)]]
agg.deaths.state[, total_deaths_rtgrowth := total_deaths/total_deaths_prev3 - 1]
agg.deaths.growth.state.plot <- ggplot(agg.deaths.state[date >= "2020-03-10"], aes(x = date, y = total_deaths_rtgrowth)) + 
  geom_line() +
    labs(x = "Date",
       y = "Rate of growth (rolling 3-day window)",
       title = "Rolling 3-day rate of growth: Total COVID-related deaths",
       subtitle = use.state) +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1), limits = c(0, max(agg.deaths.state$total_deaths_rtgrowth * 1.1)))
agg.deaths.growth.state.plot
```

## Tennessee

### Confirmed cases

```{r echo = F}
# adjust state name to flow through rest of code (use capitalized full state name)
use.state <- "Tennesseen"
cases.state <- state.raw[state %in% use.state]
agg.cases.state <- cases.state[order(date),
                          .(total_cases = sum(cases, na.rm = T)),
                          by = "date"]
agg.cases.state.plot <- ggplot(agg.cases.state[date >= "2020-03-01"], aes(x = date, y = total_cases)) + 
  geom_bar(stat = "identity") +
    labs(x = "Date",
       y = "Total confirmed COVID cases",
       title = "Total confirmed COVID cases",
       subtitle = use.state) +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) +
  scale_y_continuous(labels = comma)
agg.cases.state.plot
```

```{r echo = F}
use.state <- "Tennessee"
cases.state <- state.raw[state %in% use.state]
agg.cases.state <- cases.state[order(date),
                          .(total_cases = sum(cases, na.rm = T)),
                          by = "date"]
temp <- agg.cases.state
agg.cases.state[, total_cases_prev := temp$total_cases[match(agg.cases.state$date - 1, temp$date)]]
agg.cases.state[, total_new_cases := total_cases - total_cases_prev]
agg.new.cases.state.plot <- ggplot(agg.cases.state[date >= "2020-03-01"], aes(x = date, y = total_new_cases)) + 
  geom_bar(stat = "identity") +
    labs(x = "Date",
       y = "Confirmed new COVID cases",
       title = "Confirmed new COVID cases by day",
       subtitle = use.state) +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) +
  scale_y_continuous(labels = comma)
agg.new.cases.state.plot
```

```{r echo = F}
use.state <- "Tennessee"
cases.state <- state.raw[state %in% use.state]
agg.cases.state <- cases.state[order(date),
                          .(total_cases = sum(cases, na.rm = T)),
                          by = "date"]
temp <- agg.cases.state
agg.cases.state[, total_cases_prev3 := temp$total_cases[match(agg.cases.state$date - 3, temp$date)]]
agg.cases.state[, total_cases_rtgrowth := total_cases/total_cases_prev3 - 1]
agg.cases.growth.state.plot <- ggplot(agg.cases.state[date >= "2020-03-01"], aes(x = date, y = total_cases_rtgrowth)) + 
  geom_line() +
    labs(x = "Date",
       y = "Rate of growth (rolling 3-day window)",
       title = "Rolling 3-day rate of growth: Total confirmed COVID cases",
       subtitle = use.state) +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1), limits = c(0, max(agg.cases.state$total_cases_rtgrowth * 1.1)))
agg.cases.growth.state.plot
```

### Deaths

```{r echo = F}
# adjust state name to flow through rest of code (use capitalized full state name)
use.state <- "Tennessee"
deaths.state <- state.raw[state %in% use.state]
agg.deaths.state <- cases.state[order(date),
                          .(total_deaths = sum(deaths, na.rm = T)),
                          by = "date"]
agg.deaths.state.plot <- ggplot(agg.deaths.state[date >= "2020-03-01"], aes(x = date, y = total_deaths)) + 
  geom_bar(stat = "identity") +
    labs(x = "Date",
       y = "Total COVID-related deaths",
       title = "Total COVID-related deaths",
       subtitle = use.state) +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))
agg.deaths.state.plot
```

```{r echo = F}
use.state <- "Tennessee"
deaths.state <- state.raw[state %in% use.state]
agg.deaths.state <- deaths.state[order(date),
                          .(total_deaths = sum(deaths, na.rm = T)),
                          by = "date"]
temp <- agg.deaths.state
agg.deaths.state[, total_deaths_prev := temp$total_deaths[match(agg.deaths.state$date - 1, temp$date)]]
agg.deaths.state[, total_new_deaths := total_deaths - total_deaths_prev]
agg.new.deaths.state.plot <- ggplot(agg.deaths.state[date >= "2020-03-01"], aes(x = date, y = total_new_deaths)) + 
  geom_bar(stat = "identity") +
    labs(x = "Date",
       y = "New COVID-related deaths",
       title = "New COVID-related deaths by day",
       subtitle = use.state) +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))
agg.new.deaths.state.plot
```

```{r echo = F}
use.state <- "Tennessee"
deaths.state <- state.raw[state %in% use.state]
agg.deaths.state <- deaths.state[order(date),
                          .(total_deaths = sum(deaths, na.rm = T)),
                          by = "date"]
temp <- agg.deaths.state
agg.deaths.state[, total_deaths_prev3 := temp$total_deaths[match(agg.deaths.state$date - 3, temp$date)]]
agg.deaths.state[, total_deaths_rtgrowth := total_deaths/total_deaths_prev3 - 1]
agg.deaths.growth.state.plot <- ggplot(agg.deaths.state[date >= "2020-03-10"], aes(x = date, y = total_deaths_rtgrowth)) + 
  geom_line() +
    labs(x = "Date",
       y = "Rate of growth (rolling 3-day window)",
       title = "Rolling 3-day rate of growth: Total COVID-related deaths",
       subtitle = use.state) +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1), limits = c(0, max(agg.deaths.state$total_deaths_rtgrowth * 1.1)))
agg.deaths.growth.state.plot
```

## Pierce County, Washington

### Confirmed cases

```{r echo = F}
# adjust state and county name to flow through rest of code (use capitalized full state name)
use.state <- "Washington"
use.county <- "Pierce"
cases.county <- county.raw[state %in% use.state & county %in% use.county]
agg.cases.county <- cases.county[order(date),
                          .(total_cases = sum(cases, na.rm = T)),
                          by = "date"]
agg.cases.county.plot <- ggplot(agg.cases.county[date >= "2020-03-01"], aes(x = date, y = total_cases)) + 
  geom_bar(stat = "identity") +
    labs(x = "Date",
       y = "Total confirmed COVID cases",
       title = "Total confirmed COVID cases",
       subtitle = paste0(use.county, " County, ", use.state)) +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) +
  scale_y_continuous(labels = comma)
agg.cases.county.plot
```

```{r  echo = F, message = F, warning = F}
use.state <- "Washington"
use.county <- "Pierce"
cases.county <- county.raw[state %in% use.state & county %in% use.county]
agg.cases.county <- cases.county[order(date),
                          .(total_cases = sum(cases, na.rm = T)),
                          by = "date"]
temp <- agg.cases.county
agg.cases.county[, total_cases_prev := temp$total_cases[match(agg.cases.county$date - 1, temp$date)]]
agg.cases.county[, total_new_cases := total_cases - total_cases_prev]
agg.new.cases.county.plot <- ggplot(agg.cases.county[date >= "2020-03-01"], aes(x = date, y = total_new_cases)) + 
  geom_bar(stat = "identity") +
    labs(x = "Date",
       y = "Confirmed new COVID cases",
       title = "Confirmed new COVID cases by day",
       subtitle = paste0(use.county, " County, ", use.state)) +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) +
  scale_y_continuous(labels = comma)
agg.new.cases.county.plot
```

### Deaths

```{r echo = F}
# adjust state and county name to flow through rest of code (use capitalized full state name)
use.state <- "Washington"
use.county <- "Pierce"
deaths.county <- county.raw[state %in% use.state & county %in% use.county]
agg.deaths.county <- deaths.county[order(date),
                          .(total_deaths = sum(deaths, na.rm = T)),
                          by = "date"]
agg.deaths.county.plot <- ggplot(agg.deaths.county[date >= "2020-03-01"], aes(x = date, y = total_deaths)) + 
  geom_bar(stat = "identity") +
    labs(x = "Date",
       y = "Total COVID-related deaths",
       title = "Total COVID-related deaths",
       subtitle = paste0(use.county, " County, ", use.state)) +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))
agg.deaths.county.plot
```

```{r echo = F, message = F, warning = F}
use.state <- "Washington"
use.county <- "Pierce"
deaths.county <- county.raw[state %in% use.state & county %in% use.county]
agg.deaths.county <- deaths.county[order(date),
                          .(total_deaths = sum(deaths, na.rm = T)),
                          by = "date"]
temp <- agg.deaths.county
agg.deaths.county[, total_deaths_prev := temp$total_deaths[match(agg.deaths.county$date - 1, temp$date)]]
agg.deaths.county[, total_new_deaths := total_deaths - total_deaths_prev]
agg.new.deaths.county.plot <- ggplot(agg.deaths.county[date >= "2020-03-01"], aes(x = date, y = total_new_deaths)) + 
  geom_bar(stat = "identity") +
    labs(x = "Date",
       y = "New COVID-related deaths",
       title = "New COVID-related deaths by day",
       subtitle = paste0(use.county, " County, ", use.state)) +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))
agg.new.deaths.county.plot
```

## Orange County, California

### Confirmed cases

```{r echo = F}
# adjust state and county name to flow through rest of code (use capitalized full state name)
use.state <- "California"
use.county <- "Orange"
cases.county <- county.raw[state %in% use.state & county %in% use.county]
agg.cases.county <- cases.county[order(date),
                          .(total_cases = sum(cases, na.rm = T)),
                          by = "date"]
agg.cases.county.plot <- ggplot(agg.cases.county[date >= "2020-03-01"], aes(x = date, y = total_cases)) + 
  geom_bar(stat = "identity") +
    labs(x = "Date",
       y = "Total confirmed COVID cases",
       title = "Total confirmed COVID cases",
       subtitle = paste0(use.county, " County, ", use.state)) +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) +
  scale_y_continuous(labels = comma)
agg.cases.county.plot
```

```{r  echo = F, message = F, warning = F}
use.state <- "California"
use.county <- "Orange"
cases.county <- county.raw[state %in% use.state & county %in% use.county]
agg.cases.county <- cases.county[order(date),
                          .(total_cases = sum(cases, na.rm = T)),
                          by = "date"]
temp <- agg.cases.county
agg.cases.county[, total_cases_prev := temp$total_cases[match(agg.cases.county$date - 1, temp$date)]]
agg.cases.county[, total_new_cases := total_cases - total_cases_prev]
agg.new.cases.county.plot <- ggplot(agg.cases.county[date >= "2020-03-01"], aes(x = date, y = total_new_cases)) + 
  geom_bar(stat = "identity") +
    labs(x = "Date",
       y = "Confirmed new COVID cases",
       title = "Confirmed new COVID cases by day",
       subtitle = paste0(use.county, " County, ", use.state)) +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) +
  scale_y_continuous(labels = comma)
agg.new.cases.county.plot
```

### Deaths

```{r echo = F}
# adjust state and county name to flow through rest of code (use capitalized full state name)
use.state <- "California"
use.county <- "Orange"
deaths.county <- county.raw[state %in% use.state & county %in% use.county]
agg.deaths.county <- deaths.county[order(date),
                          .(total_deaths = sum(deaths, na.rm = T)),
                          by = "date"]
agg.deaths.county.plot <- ggplot(agg.deaths.county[date >= "2020-03-01"], aes(x = date, y = total_deaths)) + 
  geom_bar(stat = "identity") +
    labs(x = "Date",
       y = "Total COVID-related deaths",
       title = "Total COVID-related deaths",
       subtitle = paste0(use.county, " County, ", use.state)) +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))
agg.deaths.county.plot
```

```{r echo = F, message = F, warning = F}
use.state <- "California"
use.county <- "Orange"
deaths.county <- county.raw[state %in% use.state & county %in% use.county]
agg.deaths.county <- deaths.county[order(date),
                          .(total_deaths = sum(deaths, na.rm = T)),
                          by = "date"]
temp <- agg.deaths.county
agg.deaths.county[, total_deaths_prev := temp$total_deaths[match(agg.deaths.county$date - 1, temp$date)]]
agg.deaths.county[, total_new_deaths := total_deaths - total_deaths_prev]
agg.new.deaths.county.plot <- ggplot(agg.deaths.county[date >= "2020-03-01"], aes(x = date, y = total_new_deaths)) + 
  geom_bar(stat = "identity") +
    labs(x = "Date",
       y = "New COVID-related deaths",
       title = "New COVID-related deaths by day",
       subtitle = paste0(use.county, " County, ", use.state)) +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))
agg.new.deaths.county.plot
```

## Richland County, South Carolina

### Confirmed cases

```{r echo = F}
# adjust state and county name to flow through rest of code (use capitalized full state name)
use.state <- "South Carolina"
use.county <- "Richland"
cases.county <- county.raw[state %in% use.state & county %in% use.county]
agg.cases.county <- cases.county[order(date),
                          .(total_cases = sum(cases, na.rm = T)),
                          by = "date"]
agg.cases.county.plot <- ggplot(agg.cases.county[date >= "2020-03-01"], aes(x = date, y = total_cases)) + 
  geom_bar(stat = "identity") +
    labs(x = "Date",
       y = "Total confirmed COVID cases",
       title = "Total confirmed COVID cases",
       subtitle = paste0(use.county, " County, ", use.state)) +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) +
  scale_y_continuous(labels = comma)
agg.cases.county.plot
```

```{r  echo = F, message = F, warning = F}
use.state <- "South Carolina"
use.county <- "Richland"
cases.county <- county.raw[state %in% use.state & county %in% use.county]
agg.cases.county <- cases.county[order(date),
                          .(total_cases = sum(cases, na.rm = T)),
                          by = "date"]
temp <- agg.cases.county
agg.cases.county[, total_cases_prev := temp$total_cases[match(agg.cases.county$date - 1, temp$date)]]
agg.cases.county[, total_new_cases := total_cases - total_cases_prev]
agg.new.cases.county.plot <- ggplot(agg.cases.county[date >= "2020-03-01"], aes(x = date, y = total_new_cases)) + 
  geom_bar(stat = "identity") +
    labs(x = "Date",
       y = "Confirmed new COVID cases",
       title = "Confirmed new COVID cases by day",
       subtitle = paste0(use.county, " County, ", use.state)) +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) +
  scale_y_continuous(labels = comma)
agg.new.cases.county.plot
```

### Deaths

```{r echo = F}
# adjust state and county name to flow through rest of code (use capitalized full state name)
use.state <- "South Carolina"
use.county <- "Richland"
deaths.county <- county.raw[state %in% use.state & county %in% use.county]
agg.deaths.county <- deaths.county[order(date),
                          .(total_deaths = sum(deaths, na.rm = T)),
                          by = "date"]
agg.deaths.county.plot <- ggplot(agg.deaths.county[date >= "2020-03-01"], aes(x = date, y = total_deaths)) + 
  geom_bar(stat = "identity") +
    labs(x = "Date",
       y = "Total COVID-related deaths",
       title = "Total COVID-related deaths",
       subtitle = paste0(use.county, " County, ", use.state)) +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))
agg.deaths.county.plot
```

```{r echo = F, message = F, warning = F}
use.state <- "South Carolina"
use.county <- "Richland"
deaths.county <- county.raw[state %in% use.state & county %in% use.county]
agg.deaths.county <- deaths.county[order(date),
                          .(total_deaths = sum(deaths, na.rm = T)),
                          by = "date"]
temp <- agg.deaths.county
agg.deaths.county[, total_deaths_prev := temp$total_deaths[match(agg.deaths.county$date - 1, temp$date)]]
agg.deaths.county[, total_new_deaths := total_deaths - total_deaths_prev]
agg.new.deaths.county.plot <- ggplot(agg.deaths.county[date >= "2020-03-01"], aes(x = date, y = total_new_deaths)) + 
  geom_bar(stat = "identity") +
    labs(x = "Date",
       y = "New COVID-related deaths",
       title = "New COVID-related deaths by day",
       subtitle = paste0(use.county, " County, ", use.state)) +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))
agg.new.deaths.county.plot
```

## York County, South Carolina

### Confirmed cases

```{r echo = F}
# adjust state and county name to flow through rest of code (use capitalized full state name)
use.state <- "South Carolina"
use.county <- "York"
cases.county <- county.raw[state %in% use.state & county %in% use.county]
agg.cases.county <- cases.county[order(date),
                          .(total_cases = sum(cases, na.rm = T)),
                          by = "date"]
agg.cases.county.plot <- ggplot(agg.cases.county[date >= "2020-03-01"], aes(x = date, y = total_cases)) + 
  geom_bar(stat = "identity") +
    labs(x = "Date",
       y = "Total confirmed COVID cases",
       title = "Total confirmed COVID cases",
       subtitle = paste0(use.county, " County, ", use.state)) +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) +
  scale_y_continuous(labels = comma)
agg.cases.county.plot
```

```{r  echo = F, message = F, warning = F}
use.state <- "South Carolina"
use.county <- "York"
cases.county <- county.raw[state %in% use.state & county %in% use.county]
agg.cases.county <- cases.county[order(date),
                          .(total_cases = sum(cases, na.rm = T)),
                          by = "date"]
temp <- agg.cases.county
agg.cases.county[, total_cases_prev := temp$total_cases[match(agg.cases.county$date - 1, temp$date)]]
agg.cases.county[, total_new_cases := total_cases - total_cases_prev]
agg.new.cases.county.plot <- ggplot(agg.cases.county[date >= "2020-03-01"], aes(x = date, y = total_new_cases)) + 
  geom_bar(stat = "identity") +
    labs(x = "Date",
       y = "Confirmed new COVID cases",
       title = "Confirmed new COVID cases by day",
       subtitle = paste0(use.county, " County, ", use.state)) +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) +
  scale_y_continuous(labels = comma)
agg.new.cases.county.plot
```

### Deaths

```{r echo = F}
# adjust state and county name to flow through rest of code (use capitalized full state name)
use.state <- "South Carolina"
use.county <- "York"
deaths.county <- county.raw[state %in% use.state & county %in% use.county]
agg.deaths.county <- deaths.county[order(date),
                          .(total_deaths = sum(deaths, na.rm = T)),
                          by = "date"]
agg.deaths.county.plot <- ggplot(agg.deaths.county[date >= "2020-03-01"], aes(x = date, y = total_deaths)) + 
  geom_bar(stat = "identity") +
    labs(x = "Date",
       y = "Total COVID-related deaths",
       title = "Total COVID-related deaths",
       subtitle = paste0(use.county, " County, ", use.state)) +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))
agg.deaths.county.plot
```

```{r echo = F, message = F, warning = F}
use.state <- "South Carolina"
use.county <- "York"
deaths.county <- county.raw[state %in% use.state & county %in% use.county]
agg.deaths.county <- deaths.county[order(date),
                          .(total_deaths = sum(deaths, na.rm = T)),
                          by = "date"]
temp <- agg.deaths.county
agg.deaths.county[, total_deaths_prev := temp$total_deaths[match(agg.deaths.county$date - 1, temp$date)]]
agg.deaths.county[, total_new_deaths := total_deaths - total_deaths_prev]
agg.new.deaths.county.plot <- ggplot(agg.deaths.county[date >= "2020-03-01"], aes(x = date, y = total_new_deaths)) + 
  geom_bar(stat = "identity") +
    labs(x = "Date",
       y = "New COVID-related deaths",
       title = "New COVID-related deaths by day",
       subtitle = paste0(use.county, " County, ", use.state)) +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))
agg.new.deaths.county.plot
```

## Davidson County, Tennessee

### Confirmed cases

```{r echo = F}
# adjust state and county name to flow through rest of code (use capitalized full state name)
use.state <- "Tennessee"
use.county <- "Davidson"
cases.county <- county.raw[state %in% use.state & county %in% use.county]
agg.cases.county <- cases.county[order(date),
                          .(total_cases = sum(cases, na.rm = T)),
                          by = "date"]
agg.cases.county.plot <- ggplot(agg.cases.county[date >= "2020-03-01"], aes(x = date, y = total_cases)) + 
  geom_bar(stat = "identity") +
    labs(x = "Date",
       y = "Total confirmed COVID cases",
       title = "Total confirmed COVID cases",
       subtitle = paste0(use.county, " County, ", use.state)) +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) +
  scale_y_continuous(labels = comma)
agg.cases.county.plot
```

```{r  echo = F, message = F, warning = F}
use.state <- "Tennessee"
use.county <- "Davidson"
cases.county <- county.raw[state %in% use.state & county %in% use.county]
agg.cases.county <- cases.county[order(date),
                          .(total_cases = sum(cases, na.rm = T)),
                          by = "date"]
temp <- agg.cases.county
agg.cases.county[, total_cases_prev := temp$total_cases[match(agg.cases.county$date - 1, temp$date)]]
agg.cases.county[, total_new_cases := total_cases - total_cases_prev]
agg.new.cases.county.plot <- ggplot(agg.cases.county[date >= "2020-03-01"], aes(x = date, y = total_new_cases)) + 
  geom_bar(stat = "identity") +
    labs(x = "Date",
       y = "Confirmed new COVID cases",
       title = "Confirmed new COVID cases by day",
       subtitle = paste0(use.county, " County, ", use.state)) +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) +
  scale_y_continuous(labels = comma)
agg.new.cases.county.plot
```

### Deaths

```{r echo = F}
# adjust state and county name to flow through rest of code (use capitalized full state name)
use.state <- "Tennessee"
use.county <- "Davidson"
deaths.county <- county.raw[state %in% use.state & county %in% use.county]
agg.deaths.county <- deaths.county[order(date),
                          .(total_deaths = sum(deaths, na.rm = T)),
                          by = "date"]
agg.deaths.county.plot <- ggplot(agg.deaths.county[date >= "2020-03-01"], aes(x = date, y = total_deaths)) + 
  geom_bar(stat = "identity") +
    labs(x = "Date",
       y = "Total COVID-related deaths",
       title = "Total COVID-related deaths",
       subtitle = paste0(use.county, " County, ", use.state)) +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))
agg.deaths.county.plot
```

```{r echo = F, message = F, warning = F}
use.state <- "Tennessee"
use.county <- "Davidson"
deaths.county <- county.raw[state %in% use.state & county %in% use.county]
agg.deaths.county <- deaths.county[order(date),
                          .(total_deaths = sum(deaths, na.rm = T)),
                          by = "date"]
temp <- agg.deaths.county
agg.deaths.county[, total_deaths_prev := temp$total_deaths[match(agg.deaths.county$date - 1, temp$date)]]
agg.deaths.county[, total_new_deaths := total_deaths - total_deaths_prev]
agg.new.deaths.county.plot <- ggplot(agg.deaths.county[date >= "2020-03-01"], aes(x = date, y = total_new_deaths)) + 
  geom_bar(stat = "identity") +
    labs(x = "Date",
       y = "New COVID-related deaths",
       title = "New COVID-related deaths by day",
       subtitle = paste0(use.county, " County, ", use.state)) +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))
agg.new.deaths.county.plot
```

## Compare state data based on population

### Total confirmed cases

Table of total confirmed cases per million residents (all 50 states)
```{r echo = F}
state.per.cap <- merge(state.raw, stay_dates, by = "state", all.x = T)
state.per.cap[, casesPerMillion := as.integer(cases/population * 1000000)]
date.cut <- max(state.per.cap$date)
state.per.cap.agg <- state.per.cap[date == date.cut]
state.per.cap.agg <- state.per.cap.agg[order(-casesPerMillion), c("state", "casesPerMillion")]
state.per.cap.agg[, casesPerMillion := comma(casesPerMillion)]
state.per.cap.agg
```

```{r echo = F}
state.per.cap.agg[, row := .I]
top.ten <- state.per.cap.agg[row < 11, "state"]
state.per.cap.series <- state.per.cap[state %in% top.ten$state]
state.per.cap.plot <- ggplot(state.per.cap.series[date >= "2020-03-01"], aes(x = date, y = casesPerMillion, colour = state)) +
  geom_line(aes(group = state)) +
    labs(x = "Date",
       y = "Total confirmed COVID cases per million residents",
       title = "Total confirmed COVID cases per million residents",
       subtitle = paste0("States with 10 highest counts as of ", format(date.cut, "%B %d, %Y"))) +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) +
  scale_y_continuous(labels = comma)
state.per.cap.plot
```

### New confirmed cases

Table of new cases per million residents: rolling 3-day average (all 50 states)
```{r echo = F}
state.per.cap <- merge(state.raw, stay_dates, by = "state", all.x = T)
date.cut <- max(state.per.cap$date)
date.min <- max(state.per.cap$date) - 3
state.per.cap.agg <- state.per.cap[date == date.cut]
temp <- state.per.cap[date == date.min]
state.per.cap.agg[, casesPrev3 := temp$cases[match(state.per.cap.agg$state, temp$state)]]
state.per.cap.agg[, newCasesPrev3 := (cases - casesPrev3)]
state.per.cap.agg[, casesNewPerMillion := as.integer(newCasesPrev3/population * 1000000)]
state.per.cap.agg <- state.per.cap.agg[order(-casesNewPerMillion), c("state", "casesNewPerMillion")]
state.per.cap.agg[, casesNewPerMillion := comma(casesNewPerMillion)]
state.per.cap.agg
```

```{r  echo = F, message = F, warning = F}
state.per.cap.agg[, row := .I]
top.ten <- state.per.cap.agg[row < 11, "state"]
state.per.cap.series <- state.per.cap[state %in% top.ten$state]
temp <- state.per.cap.series[, c("state", "date", "cases")]
setnames(temp, "cases", "casesPrev3")
state.per.cap.series[, datePrev3 := date - 3]
state.per.cap.series <- merge(state.per.cap.series, temp, all.x = T, by.x = c("state", "datePrev3"), by.y = c("state", "date"))
state.per.cap.series[, newCasesPrev3 := (cases - casesPrev3)]
state.per.cap.series[, casesNewPerMillion := as.integer(newCasesPrev3/population * 1000000)]
state.per.cap.plot <- ggplot(state.per.cap.series[date >= "2020-03-10"], aes(x = date, y = casesNewPerMillion, colour = state)) +
  geom_line(aes(group = state)) +
    labs(x = "Date",
       y = "New confirmed COVID cases per million residents",
       title = "New confirmed COVID cases per million resident (3-day rolling average)",
       subtitle = paste0("States with 10 highest counts as of ", format(date.cut, "%B %d, %Y"))) +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) +
  scale_y_continuous(labels = comma)
state.per.cap.plot
```

### Total deaths

Table of total deaths per million residents (all 50 states)
```{r echo = F}
state.per.cap <- merge(state.raw, stay_dates, by = "state", all.x = T)
state.per.cap[, deathsPerMillion := as.integer(deaths/population * 1000000)]
date.cut <- max(state.per.cap$date)
state.per.cap.agg <- state.per.cap[date == date.cut]
state.per.cap.agg <- state.per.cap.agg[order(-deathsPerMillion), c("state", "deathsPerMillion")]
state.per.cap.agg[, deathsPerMillion := comma(deathsPerMillion)]
state.per.cap.agg
```

```{r echo = F}
state.per.cap.agg[, row := .I]
top.ten <- state.per.cap.agg[row < 11, "state"]
state.per.cap.series <- state.per.cap[state %in% top.ten$state]
state.per.cap.plot <- ggplot(state.per.cap.series[date >= "2020-03-01"], aes(x = date, y = deathsPerMillion, colour = state)) +
  geom_line(aes(group = state)) +
    labs(x = "Date",
       y = "Total COVID-related deaths per million residents",
       title = "Total COVID-related deaths per million residents",
       subtitle = paste0("States with 10 highest counts as of ", format(date.cut, "%B %d, %Y"))) +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) +
  scale_y_continuous(labels = comma)
state.per.cap.plot
```

### New deaths

Table of new deaths per million residents: rolling 3-day average (all 50 states)
```{r echo = F}
state.per.cap <- merge(state.raw, stay_dates, by = "state", all.x = T)
date.cut <- max(state.per.cap$date)
date.min <- max(state.per.cap$date) - 3
state.per.cap.agg <- state.per.cap[date == date.cut]
temp <- state.per.cap[date == date.min]
state.per.cap.agg[, deathsPrev3 := temp$deaths[match(state.per.cap.agg$state, temp$state)]]
state.per.cap.agg[, newDeathsPrev3 := (deaths - deathsPrev3)]
state.per.cap.agg[, deathsNewPerMillion := as.integer(newDeathsPrev3/population * 1000000)]
state.per.cap.agg <- state.per.cap.agg[order(-deathsNewPerMillion), c("state", "deathsNewPerMillion")]
state.per.cap.agg[, deathsNewPerMillion := comma(deathsNewPerMillion)]
state.per.cap.agg
```

```{r echo = F, message = F, warning = F}
state.per.cap.agg[, row := .I]
top.ten <- state.per.cap.agg[row < 11, "state"]
state.per.cap.series <- state.per.cap[state %in% top.ten$state]
temp <- state.per.cap.series[, c("state", "date", "deaths")]
setnames(temp, "deaths", "deathsPrev3")
state.per.cap.series[, datePrev3 := date - 3]
state.per.cap.series <- merge(state.per.cap.series, temp, all.x = T, by.x = c("state", "datePrev3"), by.y = c("state", "date"))
state.per.cap.series[, newDeathsPrev3 := (deaths - deathsPrev3)]
state.per.cap.series[, deathsNewPerMillion := as.integer(newDeathsPrev3/population * 1000000)]
state.per.cap.plot <- ggplot(state.per.cap.series[date >= "2020-03-10"], aes(x = date, y = deathsNewPerMillion, colour = state)) +
  geom_line(aes(group = state)) +
    labs(x = "Date",
       y = "New COVID-related deaths per million residents",
       title = "New COVID-related deaths per million resident (3-day rolling average)",
       subtitle = paste0("States with 10 highest counts as of ", format(date.cut, "%B %d, %Y"))) +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))
state.per.cap.plot
```

## Compare new cases based on stay-at-home orders

```{r echo = F}
state.dt <- state.raw
state.dt[, stay_date := stay_dates$start_date[match(state.raw$state, stay_dates$state)]]
state.dt[, population := stay_dates$population[match(state.raw$state, stay_dates$state)]]
#state.dt[, flgOrder := ifelse(is.na(stay_date), 0, 1)]
state.dt[, flgOrder := ifelse(is.na(stay_date) | stay_date >= date, "No", "Yes")]
#state.dt[, ctDaysFromBaseline := ifelse(flgOrder == 0, as.integer(date - as.Date("2020-03-10")),
#                                         as.integer(date - stay_date))]
#state.dt[, ctDaysFromBaseline := as.integer(date - stay_date)]
temp <- state.dt
temp <- temp[, c("date", "state", "cases")]
setnames(temp, "cases", "cases_prev3")
state.dt[, dtPrev3 := date - 3]
state.dt <- merge(state.raw, temp, all.x = T, by.x = c("state", "dtPrev3"), by.y = c("state", "date"))
state.dt[, casesNewPerMillion := as.integer((cases - cases_prev3)/population * 1000000)]
state.agg.comp <- state.dt[date >= "2020-03-10" & !state %in% "Puerto Rico"]
state.agg.comp <- state.agg.comp[order(date, flgOrder),
                          .(avgNewCases = mean(casesNewPerMillion, na.rm = T),
                            cts = .N),
                           by = c("date", "flgOrder")]
state.agg.comp.plot <- ggplot(state.agg.comp, aes(x = date, y = avgNewCases, colour = as.factor(flgOrder))) +
  geom_point(aes(size = cts)) +
  geom_line(aes(group = flgOrder)) +
    labs(x = "Date",
       y = "New confirmed COVID cases per million residents",
       title = "New COVID cases vs.stay-at-home order",
       subtitle = "New cases are a 3-day rolling average per million residents",
       size = "Count of states",
       colour = "Stay-at-home order") +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))
state.agg.comp.plot
```
