---
title: "Covid plots"
output: html_notebook
---

## Set-up
Load packages and set options
```{r include = F, results = 'hide'}
# load packages
library(data.table)
library(ggplot2)
library(scales)
library(dplyr)
library(lubridate)
library(utils)
library(httr)
library(graphics)

# turn off scientific notation
options(scipen = 999)
```

Read in files from NYT opensource database
```{r include = F, results = 'hide'}
state.raw <- data.table(read.csv(file=url("https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-states.csv"), header = T))
county.raw <- data.table(read.csv(file=url("https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-counties.csv"), header = T))
```

Add useful fields & local files
```{r include = F, results = 'hide'}
# format dates
state.raw[, date := as.Date(date)]
county.raw[, date := as.Date(date)]

# grad stay-at-home dates (manually updated locally)
stay_dates <- fread("~/covid19/raw_data/stay_at_home_dates.csv", header = T)
stay_dates[, start_date := as.Date(start_date, format = "%Y-%m-%d")]
stay_dates[, population := as.integer(gsub(",", "", population))]
```


## Total counts
Confirmed cases -- US
```{r}
agg.cases.US <- state.raw[order(date),
                          .(total_cases = sum(cases, na.rm = T)),
                          by = "date"]
agg.cases.US.plot <- ggplot(agg.cases.US[date >= "2020-03-01"], aes(x = date, y = total_cases)) + 
  geom_bar(stat = "identity") +
    labs(x = "Date",
       y = "Total confirmed COVID cases",
       title = "Total confirmed COVID cases",
       subtitle = "United States") +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) +
  scale_y_continuous(labels = comma)
agg.cases.US.plot
```

Confirmed cases -- specific state
```{r}
# adjust state name to flow through rest of code (use capitalized full state name)
use.state <- "South Carolina"
cases.state <- state.raw[state %in% use.state]
agg.cases.state <- cases.state[order(date),
                          .(total_cases = sum(cases, na.rm = T)),
                          by = "date"]
agg.cases.state.plot <- ggplot(agg.cases.state[date >= "2020-03-01"], aes(x = date, y = total_cases)) + 
  geom_bar(stat = "identity") +
    labs(x = "Date",
       y = "Total confirmed COVID cases",
       title = "Total confirmed COVID cases",
       subtitle = use.state) +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) +
  scale_y_continuous(labels = comma)
agg.cases.state.plot
```

Confirmed cases -- specific county
```{r}
# adjust state and county name to flow through rest of code (use capitalized full state name)
use.state <- "Washington"
use.county <- "Pierce"
cases.county <- county.raw[state %in% use.state & county %in% use.county]
agg.cases.county <- cases.county[order(date),
                          .(total_cases = sum(cases, na.rm = T)),
                          by = "date"]
agg.cases.county.plot <- ggplot(agg.cases.county[date >= "2020-03-01"], aes(x = date, y = total_cases)) + 
  geom_bar(stat = "identity") +
    labs(x = "Date",
       y = "Total confirmed COVID cases",
       title = "Total confirmed COVID cases",
       subtitle = paste0(use.county, " County, ", use.state)) +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) +
  scale_y_continuous(labels = comma)
agg.cases.county.plot
```

Deaths -- US
```{r}
agg.deaths.US <- state.raw[order(date),
                          .(total_deaths = sum(deaths, na.rm = T)),
                          by = "date"]
agg.deaths.US.plot <- ggplot(agg.deaths.US[date >= "2020-03-01"], aes(x = date, y = total_deaths)) + 
  geom_bar(stat = "identity") +
    labs(x = "Date",
       y = "Total COVID-related deaths",
       title = "Total COVID-related deaths",
       subtitle = "United States") +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) +
  scale_y_continuous(labels = comma)
agg.deaths.US.plot
```

Deaths -- specific state
```{r}
# adjust state name to flow through rest of code (use capitalized full state name)
use.state <- "Washington"
deaths.state <- state.raw[state %in% use.state]
agg.deaths.state <- cases.state[order(date),
                          .(total_deaths = sum(deaths, na.rm = T)),
                          by = "date"]
agg.deaths.state.plot <- ggplot(agg.deaths.state[date >= "2020-03-01"], aes(x = date, y = total_deaths)) + 
  geom_bar(stat = "identity") +
    labs(x = "Date",
       y = "Total COVID-related deaths",
       title = "Total COVID-related deaths",
       subtitle = use.state) +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))
agg.deaths.state.plot
```

Deaths -- specific county
```{r}
# adjust state and county name to flow through rest of code (use capitalized full state name)
use.state <- "Washington"
use.county <- "Pierce"
deaths.county <- county.raw[state %in% use.state & county %in% use.county]
agg.deaths.county <- deaths.county[order(date),
                          .(total_deaths = sum(deaths, na.rm = T)),
                          by = "date"]
agg.deaths.county.plot <- ggplot(agg.deaths.county[date >= "2020-03-01"], aes(x = date, y = total_deaths)) + 
  geom_bar(stat = "identity") +
    labs(x = "Date",
       y = "Total COVID-related deaths",
       title = "Total COVID-related deaths",
       subtitle = paste0(use.county, " County, ", use.state)) +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))
agg.deaths.county.plot
```

## New counts
Confirmed cases -- US
```{r}
agg.cases.US <- state.raw[order(date),
                          .(total_cases = sum(cases, na.rm = T)),
                          by = "date"]
temp <- agg.cases.US
agg.cases.US[, total_cases_prev := temp$total_cases[match(agg.cases.US$date - 1, temp$date)]]
agg.cases.US[, total_new_cases := total_cases - total_cases_prev]
agg.new.cases.US.plot <- ggplot(agg.cases.US[date >= "2020-03-01"], aes(x = date, y = total_new_cases)) + 
  geom_bar(stat = "identity") +
    labs(x = "Date",
       y = "Confirmed new COVID cases",
       title = "Confirmed new COVID cases by day",
       subtitle = "United States") +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) +
  scale_y_continuous(labels = comma)
agg.new.cases.US.plot
```

Confirmed cases -- specific state
```{r}
use.state <- "Washington"
cases.state <- state.raw[state %in% use.state]
agg.cases.state <- cases.state[order(date),
                          .(total_cases = sum(cases, na.rm = T)),
                          by = "date"]
temp <- agg.cases.state
agg.cases.state[, total_cases_prev := temp$total_cases[match(agg.cases.state$date - 1, temp$date)]]
agg.cases.state[, total_new_cases := total_cases - total_cases_prev]
agg.new.cases.state.plot <- ggplot(agg.cases.state[date >= "2020-03-01"], aes(x = date, y = total_new_cases)) + 
  geom_bar(stat = "identity") +
    labs(x = "Date",
       y = "Confirmed new COVID cases",
       title = "Confirmed new COVID cases by day",
       subtitle = use.state) +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) +
  scale_y_continuous(labels = comma)
agg.new.cases.state.plot
```

Confirmed cases -- specific county
```{r}
use.state <- "Washington"
use.county <- "Pierce"
cases.county <- county.raw[state %in% use.state & county %in% use.county]
agg.cases.county <- cases.county[order(date),
                          .(total_cases = sum(cases, na.rm = T)),
                          by = "date"]
temp <- agg.cases.county
agg.cases.county[, total_cases_prev := temp$total_cases[match(agg.cases.county$date - 1, temp$date)]]
agg.cases.county[, total_new_cases := total_cases - total_cases_prev]
agg.new.cases.county.plot <- ggplot(agg.cases.county[date >= "2020-03-01"], aes(x = date, y = total_new_cases)) + 
  geom_bar(stat = "identity") +
    labs(x = "Date",
       y = "Confirmed new COVID cases",
       title = "Confirmed new COVID cases by day",
       subtitle = paste0(use.county, " County, ", use.state)) +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) +
  scale_y_continuous(labels = comma)
agg.new.cases.county.plot
```

Deaths -- US
```{r}
agg.deaths.US <- state.raw[order(date),
                          .(total_deaths = sum(deaths, na.rm = T)),
                          by = "date"]
temp <- agg.deaths.US
agg.deaths.US[, total_deaths_prev := temp$total_deaths[match(agg.deaths.US$date - 1, temp$date)]]
agg.deaths.US[, total_new_deaths := total_deaths - total_deaths_prev]
agg.new.deaths.US.plot <- ggplot(agg.deaths.US[date >= "2020-03-01"], aes(x = date, y = total_new_deaths)) + 
  geom_bar(stat = "identity") +
    labs(x = "Date",
       y = "New COVID-related deaths",
       title = "New COVID-related deaths by day",
       subtitle = "United States") +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) +
  scale_y_continuous(labels = comma)
agg.new.deaths.US.plot
```

Deaths -- specific state
```{r}
use.state <- "Washington"
deaths.state <- state.raw[state %in% use.state]
agg.deaths.state <- deaths.state[order(date),
                          .(total_deaths = sum(deaths, na.rm = T)),
                          by = "date"]
temp <- agg.deaths.state
agg.deaths.state[, total_deaths_prev := temp$total_deaths[match(agg.deaths.state$date - 1, temp$date)]]
agg.deaths.state[, total_new_deaths := total_deaths - total_deaths_prev]
agg.new.deaths.state.plot <- ggplot(agg.deaths.state[date >= "2020-03-01"], aes(x = date, y = total_new_deaths)) + 
  geom_bar(stat = "identity") +
    labs(x = "Date",
       y = "New COVID-related deaths",
       title = "New COVID-related deaths by day",
       subtitle = use.state) +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))
agg.new.deaths.state.plot
```

Deaths -- specific county
```{r}
use.state <- "Washington"
use.county <- "Pierce"
deaths.county <- county.raw[state %in% use.state & county %in% use.county]
agg.deaths.county <- deaths.county[order(date),
                          .(total_deaths = sum(deaths, na.rm = T)),
                          by = "date"]
temp <- agg.deaths.county
agg.deaths.county[, total_deaths_prev := temp$total_deaths[match(agg.deaths.county$date - 1, temp$date)]]
agg.deaths.county[, total_new_deaths := total_deaths - total_deaths_prev]
agg.new.deaths.county.plot <- ggplot(agg.deaths.county[date >= "2020-03-01"], aes(x = date, y = total_new_deaths)) + 
  geom_bar(stat = "identity") +
    labs(x = "Date",
       y = "New COVID-related deaths",
       title = "New COVID-related deaths by day",
       subtitle = paste0(use.county, " County, ", use.state)) +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))
agg.new.deaths.county.plot
```

## Rate of growth
Confirmed cases -- US
```{r}
agg.cases.US <- state.raw[order(date),
                          .(total_cases = sum(cases, na.rm = T)),
                          by = "date"]
temp <- agg.cases.US
agg.cases.US[, total_cases_prev3 := temp$total_cases[match(agg.cases.US$date - 3, temp$date)]]
agg.cases.US[, total_cases_rtgrowth := total_cases/total_cases_prev3 - 1]
agg.cases.growth.US.plot <- ggplot(agg.cases.US[date >= "2020-03-01"], aes(x = date, y = total_cases_rtgrowth)) + 
  geom_line() +
    labs(x = "Date",
       y = "Rate of growth (rolling 3-day window)",
       title = "Rolling 3-day rate of growth: Total confirmed COVID cases",
       subtitle = "United States") +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1), limits = c(0, max(agg.cases.US$total_cases_rtgrowth * 1.1)))
agg.cases.growth.US.plot
```

Confirmed cases -- specific state
```{r}
use.state <- "Washington"
cases.state <- state.raw[state %in% use.state]
agg.cases.state <- cases.state[order(date),
                          .(total_cases = sum(cases, na.rm = T)),
                          by = "date"]
temp <- agg.cases.state
agg.cases.state[, total_cases_prev3 := temp$total_cases[match(agg.cases.state$date - 3, temp$date)]]
agg.cases.state[, total_cases_rtgrowth := total_cases/total_cases_prev3 - 1]
agg.cases.growth.state.plot <- ggplot(agg.cases.state[date >= "2020-03-01"], aes(x = date, y = total_cases_rtgrowth)) + 
  geom_line() +
    labs(x = "Date",
       y = "Rate of growth (rolling 3-day window)",
       title = "Rolling 3-day rate of growth: Total confirmed COVID cases",
       subtitle = use.state) +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1), limits = c(0, max(agg.cases.state$total_cases_rtgrowth * 1.1)))
agg.cases.growth.state.plot
```

Confirmed cases -- specific county
```{r}
use.state <- "Washington"
use.county <- "Pierce"
cases.county <- county.raw[state %in% use.state & county %in% use.county]
agg.cases.county <- cases.county[order(date),
                          .(total_cases = sum(cases, na.rm = T)),
                          by = "date"]
temp <- agg.cases.county
agg.cases.county[, total_cases_prev3 := temp$total_cases[match(agg.cases.county$date - 3, temp$date)]]
agg.cases.county[, total_cases_rtgrowth := total_cases/total_cases_prev3 - 1]
agg.cases.growth.county.plot <- ggplot(agg.cases.county[date >= "2020-03-01"], aes(x = date, y = total_cases_rtgrowth)) + 
  geom_line() +
    labs(x = "Date",
       y = "Rate of growth (rolling 3-day window)",
       title = "Rolling 3-day rate of growth: Total confirmed COVID cases",
       subtitle = paste0(use.county, " County, ", use.state)) +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1), limits = c(0, max(agg.cases.county$total_cases_rtgrowth * 1.1)))
agg.cases.growth.county.plot
```

Deaths -- US
```{r}
agg.deaths.US <- state.raw[order(date),
                          .(total_deaths = sum(deaths, na.rm = T)),
                          by = "date"]
temp <- agg.deaths.US
agg.deaths.US[, total_deaths_prev3 := temp$total_deaths[match(agg.deaths.US$date - 3, temp$date)]]
agg.deaths.US[, total_deaths_rtgrowth := total_deaths/total_deaths_prev3 - 1]
agg.deaths.growth.US.plot <- ggplot(agg.deaths.US[date >= "2020-03-10"], aes(x = date, y = total_deaths_rtgrowth)) + 
  geom_line() +
    labs(x = "Date",
       y = "Rate of growth (rolling 3-day window)",
       title = "Rolling 3-day rate of growth: Total COVID-related deaths",
       subtitle = "United States") +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1), limits = c(0, max(agg.deaths.US$total_deaths_rtgrowth * 1.1)))
agg.deaths.growth.US.plot
```

Deaths -- specific state
```{r}
use.state <- "Washington"
deaths.state <- state.raw[state %in% use.state]
agg.deaths.state <- deaths.state[order(date),
                          .(total_deaths = sum(deaths, na.rm = T)),
                          by = "date"]
temp <- agg.deaths.state
agg.deaths.state[, total_deaths_prev3 := temp$total_deaths[match(agg.deaths.state$date - 3, temp$date)]]
agg.deaths.state[, total_deaths_rtgrowth := total_deaths/total_deaths_prev3 - 1]
agg.deaths.growth.state.plot <- ggplot(agg.deaths.state[date >= "2020-03-10"], aes(x = date, y = total_deaths_rtgrowth)) + 
  geom_line() +
    labs(x = "Date",
       y = "Rate of growth (rolling 3-day window)",
       title = "Rolling 3-day rate of growth: Total COVID-related deaths",
       subtitle = use.state) +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1), limits = c(0, max(agg.deaths.state$total_deaths_rtgrowth * 1.1)))
agg.deaths.growth.state.plot
```

## Examine state figures per capita
Table of total confirmed cases per capita (all states)
```{r}
state.per.cap <- merge(state.raw, stay_dates, by = "state", all.x = T)
state.per.cap[, casesPerMillion := as.integer(cases/population * 1000000)]
date.cut <- max(state.per.cap$date)
state.per.cap.agg <- state.per.cap[date == date.cut]
state.per.cap.agg <- state.per.cap.agg[order(-casesPerMillion), c("state", "casesPerMillion")]
state.per.cap.agg[, casesPerMillion := comma(casesPerMillion)]
state.per.cap.agg
```

Plot of total confirmed cases per capita over time (top 10 states as of most recent day's data)
```{r}
state.per.cap.agg[, row := .I]
top.ten <- state.per.cap.agg[row < 11, "state"]
state.per.cap.series <- state.per.cap[state %in% top.ten$state]
state.per.cap.plot <- ggplot(state.per.cap.series[date >= "2020-03-01"], aes(x = date, y = casesPerMillion, colour = state)) +
  geom_line(aes(group = state)) +
    labs(x = "Date",
       y = "Total confirmed COVID cases per million residents",
       title = "Total confirmed COVID cases per million residents",
       subtitle = paste0("States with 10 highest counts as of ", format(date.cut, "%B %d, %Y"))) +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) +
  scale_y_continuous(labels = comma)
state.per.cap.plot
```

Table of new cases per capita: rolling 3-day average (all 50 states)
```{r}
state.per.cap <- merge(state.raw, stay_dates, by = "state", all.x = T)
date.cut <- max(state.per.cap$date)
date.min <- max(state.per.cap$date) - 3
state.per.cap.agg <- state.per.cap[date == date.cut]
temp <- state.per.cap[date == date.min]
state.per.cap.agg[, casesPrev3 := temp$cases[match(state.per.cap.agg$state, temp$state)]]
state.per.cap.agg[, newCasesPrev3 := (cases - casesPrev3)]
state.per.cap.agg[, casesNewPerMillion := as.integer(newCasesPrev3/population * 1000000)]
state.per.cap.agg <- state.per.cap.agg[order(-casesNewPerMillion), c("state", "casesNewPerMillion")]
state.per.cap.agg[, casesNewPerMillion := comma(casesNewPerMillion)]
state.per.cap.agg
```

Plot of new cases per capita: rolling 3-day total (top 10 states as of most recent day's data)
```{r}
state.per.cap.agg[, row := .I]
top.ten <- state.per.cap.agg[row < 11, "state"]
state.per.cap.series <- state.per.cap[state %in% top.ten$state]
temp <- state.per.cap.series[, c("state", "date", "cases")]
setnames(temp, "cases", "casesPrev3")
state.per.cap.series[, datePrev3 := date - 3]
state.per.cap.series <- merge(state.per.cap.series, temp, all.x = T, by.x = c("state", "datePrev3"), by.y = c("state", "date"))
state.per.cap.series[, newCasesPrev3 := (cases - casesPrev3)]
state.per.cap.series[, casesNewPerMillion := as.integer(newCasesPrev3/population * 1000000)]
state.per.cap.plot <- ggplot(state.per.cap.series[date >= "2020-03-10"], aes(x = date, y = casesNewPerMillion, colour = state)) +
  geom_line(aes(group = state)) +
    labs(x = "Date",
       y = "New confirmed COVID cases per million residents",
       title = "New confirmed COVID cases per million resident (3-day rolling average)",
       subtitle = paste0("States with 10 highest counts as of ", format(date.cut, "%B %d, %Y"))) +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) +
  scale_y_continuous(labels = comma)
state.per.cap.plot
```

Table of total deaths per capita (all states)
```{r}
state.per.cap <- merge(state.raw, stay_dates, by = "state", all.x = T)
state.per.cap[, deathsPerMillion := as.integer(deaths/population * 1000000)]
date.cut <- max(state.per.cap$date)
state.per.cap.agg <- state.per.cap[date == date.cut]
state.per.cap.agg <- state.per.cap.agg[order(-deathsPerMillion), c("state", "deathsPerMillion")]
state.per.cap.agg[, deathsPerMillion := comma(deathsPerMillion)]
state.per.cap.agg
```

Plot of total deaths per capita over time (top 10 states as of most recent day's data)
```{r}
state.per.cap.agg[, row := .I]
top.ten <- state.per.cap.agg[row < 11, "state"]
state.per.cap.series <- state.per.cap[state %in% top.ten$state]
state.per.cap.plot <- ggplot(state.per.cap.series[date >= "2020-03-01"], aes(x = date, y = deathsPerMillion, colour = state)) +
  geom_line(aes(group = state)) +
    labs(x = "Date",
       y = "Total COVID-related deaths per million residents",
       title = "Total COVID-related deaths per million residents",
       subtitle = paste0("States with 10 highest counts as of ", format(date.cut, "%B %d, %Y"))) +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) +
  scale_y_continuous(labels = comma)
state.per.cap.plot
```

Table of new deaths per capita: rolling 3-day average (all 50 states)
```{r}
state.per.cap <- merge(state.raw, stay_dates, by = "state", all.x = T)
date.cut <- max(state.per.cap$date)
date.min <- max(state.per.cap$date) - 3
state.per.cap.agg <- state.per.cap[date == date.cut]
temp <- state.per.cap[date == date.min]
state.per.cap.agg[, deathsPrev3 := temp$deaths[match(state.per.cap.agg$state, temp$state)]]
state.per.cap.agg[, newDeathsPrev3 := (deaths - deathsPrev3)]
state.per.cap.agg[, deathsNewPerMillion := as.integer(newDeathsPrev3/population * 1000000)]
state.per.cap.agg <- state.per.cap.agg[order(-deathsNewPerMillion), c("state", "deathsNewPerMillion")]
state.per.cap.agg[, deathsNewPerMillion := comma(deathsNewPerMillion)]
state.per.cap.agg
```

Plot of new deaths per capita over time: rolling 3-day total (top 10 states as of most recent day's data)
```{r}
state.per.cap.agg[, row := .I]
top.ten <- state.per.cap.agg[row < 11, "state"]
state.per.cap.series <- state.per.cap[state %in% top.ten$state]
temp <- state.per.cap.series[, c("state", "date", "deaths")]
setnames(temp, "deaths", "deathsPrev3")
state.per.cap.series[, datePrev3 := date - 3]
state.per.cap.series <- merge(state.per.cap.series, temp, all.x = T, by.x = c("state", "datePrev3"), by.y = c("state", "date"))
state.per.cap.series[, newDeathsPrev3 := (deaths - deathsPrev3)]
state.per.cap.series[, deathsNewPerMillion := as.integer(newDeathsPrev3/population * 1000000)]
state.per.cap.plot <- ggplot(state.per.cap.series[date >= "2020-03-10"], aes(x = date, y = deathsNewPerMillion, colour = state)) +
  geom_line(aes(group = state)) +
    labs(x = "Date",
       y = "New COVID-related deaths per million residents",
       title = "New COVID-related deaths per million resident (3-day rolling average)",
       subtitle = paste0("States with 10 highest counts as of ", format(date.cut, "%B %d, %Y"))) +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))
state.per.cap.plot
```

## Explore growth rates based on stay-at-home orders
Confirmed cases
```{r}
state.dt <- state.raw
state.dt[, stay_date := stay_dates$start_date[match(state.raw$state, stay_dates$state)]]
state.dt[, population := stay_dates$population[match(state.raw$state, stay_dates$state)]]
#state.dt[, flgOrder := ifelse(is.na(stay_date), 0, 1)]
state.dt[, flgOrder := ifelse(is.na(stay_date) | stay_date >= date, "No", "Yes")]
#state.dt[, ctDaysFromBaseline := ifelse(flgOrder == 0, as.integer(date - as.Date("2020-03-10")),
#                                         as.integer(date - stay_date))]
#state.dt[, ctDaysFromBaseline := as.integer(date - stay_date)]
temp <- state.dt
temp <- temp[, c("date", "state", "cases")]
setnames(temp, "cases", "cases_prev3")
state.dt[, dtPrev3 := date - 3]
state.dt <- merge(state.raw, temp, all.x = T, by.x = c("state", "dtPrev3"), by.y = c("state", "date"))
state.dt[, casesNewPerMillion := as.integer((cases - cases_prev3)/population * 1000000)]
state.agg.comp <- state.dt[date >= "2020-03-10" & !state %in% "Puerto Rico"]
state.agg.comp <- state.agg.comp[order(date, flgOrder),
                          .(avgNewCases = mean(casesNewPerMillion, na.rm = T),
                            cts = .N),
                           by = c("date", "flgOrder")]
state.agg.comp.plot <- ggplot(state.agg.comp, aes(x = date, y = avgNewCases, colour = as.factor(flgOrder))) +
  geom_point(aes(size = cts)) +
  geom_line(aes(group = flgOrder)) +
    labs(x = "Date",
       y = "New confirmed COVID cases per million residents",
       title = "Avg new COVID cases in states with and without stay-at-home order",
       subtitle = "New cases are a 3-day rolling average per million residents",
       size = "Count of states",
       colour = "Stay-at-home order for state") +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))
state.agg.comp.plot
```
