---
title: "Covid plots for family"
output:
  html_document:
    toc: true
    toc_depth: 4
    fig_width: 10
    code_folding: "hide"
date: '`r format(Sys.time(), "%B %d, %Y")`'
---

## Background
Notes and format last updated Apr 9, 2020

* This page shows COVID data broken down into regions that affect our family. It pulls from publicly available raw data from the NY Times, which is reported daily. It's on a one-day lag, so when you see this report, the most recent day included will be the day prior.

* The page starts by showing growth rates. I've moved down the raw numbers to later in the page because you can see them everywhere, and it's sometimes hard to understand trends looking at them. Growth rates seem to me to be more important for understanding our progress.
  * For the U.S. and each of our states, we show two growth rate charts: confirmed cases (total and new) and deaths (total and new). Each chart has two lines: a growth-rate in the rolling three-day *total* of cases (or deaths) and a growth rate in a rolling three-day average of *new* cases (or deaths).
    * For total cases (or deaths), the charts show the growth rate in total cases (or deaths) over the previous three days for each respective date. For example, a three-day growth rate of 100% for total cases means that total cases are doubling every three days. This rate can't go below zero (because total cases (or deaths) can never go down), but we want it to get as close to zero as possible.
    * For new cases (or deaths), the charts show the daily growth rate in the average number of new cases (or deaths) for the previous three days. This rate can go below zero because we'll eventually start to see consistently less new cases (or deaths) each day. When the new case (or death) growth rate is consistently below zero, it means we're on the other side of the "apex."

* The page then shows state and county rankings normalized for population (number of cases or deaths per million residents).
  * For states, it shows tables ranking all 50 states for per populations rates of total cases, new cases, total deaths, and new deaths. In addition to the tables, the trends for the top states are plotted over time.
  * For counties, it shows tables with our counties ranked by percentile of U.S. counties for per population rates of total cases and total deaths. The table also shows the top counties in the country for perspecive. And there is then a plot showing our counties' percentiles over time.

* There is then a long section with raw counts for the U.S., our states, and our counties for each of these categories:
  * Total confirmed cases: This is the cumulative total of cases reported.
  * New confirmed cases: This is the number of new cases reported on each respective date.
  * Total deaths: This is the cumulative total of deaths reported.
  * New deaths: This is the number of new deaths reported on each respective date.
  * Rolling three-day death growth rate: This is the same as described for total and new cases above, but for total and new deaths.

* Finally, the last section compares data from states that have and have not enacted state-wide stay-at-home orders, trying to see if there are any effects of the orders. It's a work in progress, and there are more extensive notes in that section.

```{r include = F, results = 'hide'}
# load packages
library(data.table)
library(ggplot2)
library(scales)
library(dplyr)
library(lubridate)
library(utils)
library(httr)
library(graphics)

# turn off scientific notation
options(scipen = 999)

```

```{r include = F, results = 'hide'}
state.raw <- data.table(read.csv(file=url("https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-states.csv"), header = T))
county.raw <- data.table(read.csv(file=url("https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-counties.csv"), header = T))
```

```{r include = F, results = 'hide'}
# format dates
state.raw[, date := as.Date(date)]
county.raw[, date := as.Date(date)]

# stay-at-home dates (manually updated locally)
stay_dates <- fread("~/covid19/raw_data/stay_at_home_dates.csv", header = T)
stay_dates[, start_date := as.Date(start_date, format = "%Y-%m-%d")]
stay_dates[, population := as.integer(gsub(",", "", population))]

# county populations
county_pop <- fread("~/covid19/raw_data/county_pop_data.csv", header = T)
# county_pop <- county_pop[grep("County", CTYNAME)]
county_pop[, CTYNAME := gsub("(.*) County", "\\1", CTYNAME)]
```

## Growth rates

### Confirmed cases (total and new)

#### U.S.

```{r message = F, warning = F}
agg.cases.US <- state.raw[order(date),
                          .(total_cases = sum(cases, na.rm = T)),
                          by = "date"]

temp <- agg.cases.US
agg.cases.US[, total_cases_prev3 := temp$total_cases[match(agg.cases.US$date - 3, temp$date)]]
agg.cases.US[, total_cases_rtgrowth := total_cases/total_cases_prev3 - 1]
agg.cases.US[, new_cases_prev3 := total_cases - total_cases_prev3]

temp <- agg.cases.US
agg.cases.US[, new_cases_prev3_last := temp$new_cases_prev3[match(agg.cases.US$date - 1, temp$date)]]
agg.cases.US[, new_cases_rtgrowth := new_cases_prev3/new_cases_prev3_last - 1]

melt.use <- melt(agg.cases.US[, c("date", "total_cases_rtgrowth", "new_cases_rtgrowth")], id.vars = "date")
melt.use[, variable := ifelse(variable %in% "total_cases_rtgrowth", "Total Cases", "New Cases")]

agg.cases.growth.US.plot <- ggplot(melt.use[date >= "2020-03-16"], aes(x = date, y = value, color = variable)) + 
  geom_line(aes(group = variable)) +
    labs(x = "Date",
       y = "Rate of growth (rolling 3-day window)",
       title = "Rolling 3-day rate of growth: Confirmed COVID cases",
       subtitle = "United States",
       color = "Metric") +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1), limits = 
                       c(ifelse(min(melt.use$value) > 0, 0, min(melt.use$value)), max(melt.use$value * 1.1)))
agg.cases.growth.US.plot
```

#### Our states

```{r message = F, warning = F}
use.state <- "Washington"
cases.state <- state.raw[state %in% use.state]
agg.cases.state <- cases.state[order(date),
                          .(total_cases = sum(cases, na.rm = T)),
                          by = "date"]

temp <- agg.cases.state
agg.cases.state[, total_cases_prev3 := temp$total_cases[match(agg.cases.state$date - 3, temp$date)]]
agg.cases.state[, total_cases_rtgrowth := total_cases/total_cases_prev3 - 1]
agg.cases.state[, new_cases_prev3 := total_cases - total_cases_prev3]

temp <- agg.cases.state
agg.cases.state[, new_cases_prev3_last := temp$new_cases_prev3[match(agg.cases.state$date - 1, temp$date)]]
agg.cases.state[, new_cases_rtgrowth := new_cases_prev3/new_cases_prev3_last - 1]

melt.use <- melt(agg.cases.state[, c("date", "total_cases_rtgrowth", "new_cases_rtgrowth")], id.vars = "date")
melt.use[, variable := ifelse(variable %in% "total_cases_rtgrowth", "Total Cases", "New Cases")]

agg.cases.growth.state.plot <- ggplot(melt.use[date >= "2020-03-16"], aes(x = date, y = value, color = variable)) + 
  geom_line(aes(group = variable)) +
    labs(x = "Date",
       y = "Rate of growth (rolling 3-day window)",
       title = "Rolling 3-day rate of growth: Confirmed COVID cases",
       subtitle = use.state,
       color = "Metric") +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1), limits = 
                       c(ifelse(min(melt.use$value) > 0, 0, min(melt.use$value)), max(melt.use$value * 1.1)))
agg.cases.growth.state.plot
```

```{r message = F, warning = F}
use.state <- "California"
cases.state <- state.raw[state %in% use.state]
agg.cases.state <- cases.state[order(date),
                          .(total_cases = sum(cases, na.rm = T)),
                          by = "date"]

temp <- agg.cases.state
agg.cases.state[, total_cases_prev3 := temp$total_cases[match(agg.cases.state$date - 3, temp$date)]]
agg.cases.state[, total_cases_rtgrowth := total_cases/total_cases_prev3 - 1]
agg.cases.state[, new_cases_prev3 := total_cases - total_cases_prev3]

temp <- agg.cases.state
agg.cases.state[, new_cases_prev3_last := temp$new_cases_prev3[match(agg.cases.state$date - 1, temp$date)]]
agg.cases.state[, new_cases_rtgrowth := new_cases_prev3/new_cases_prev3_last - 1]

melt.use <- melt(agg.cases.state[, c("date", "total_cases_rtgrowth", "new_cases_rtgrowth")], id.vars = "date")
melt.use[, variable := ifelse(variable %in% "total_cases_rtgrowth", "Total Cases", "New Cases")]

agg.cases.growth.state.plot <- ggplot(melt.use[date >= "2020-03-16"], aes(x = date, y = value, color = variable)) + 
  geom_line(aes(group = variable)) +
    labs(x = "Date",
       y = "Rate of growth (rolling 3-day window)",
       title = "Rolling 3-day rate of growth: Confirmed COVID cases",
       subtitle = use.state,
       color = "Metric") +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1), limits = 
                       c(ifelse(min(melt.use$value) > 0, 0, min(melt.use$value)), max(melt.use$value * 1.1)))
agg.cases.growth.state.plot
```

```{r message = F, warning = F}
use.state <- "South Carolina"
cases.state <- state.raw[state %in% use.state]
agg.cases.state <- cases.state[order(date),
                          .(total_cases = sum(cases, na.rm = T)),
                          by = "date"]

temp <- agg.cases.state
agg.cases.state[, total_cases_prev3 := temp$total_cases[match(agg.cases.state$date - 3, temp$date)]]
agg.cases.state[, total_cases_rtgrowth := total_cases/total_cases_prev3 - 1]
agg.cases.state[, new_cases_prev3 := total_cases - total_cases_prev3]

temp <- agg.cases.state
agg.cases.state[, new_cases_prev3_last := temp$new_cases_prev3[match(agg.cases.state$date - 1, temp$date)]]
agg.cases.state[, new_cases_rtgrowth := new_cases_prev3/new_cases_prev3_last - 1]

melt.use <- melt(agg.cases.state[, c("date", "total_cases_rtgrowth", "new_cases_rtgrowth")], id.vars = "date")
melt.use[, variable := ifelse(variable %in% "total_cases_rtgrowth", "Total Cases", "New Cases")]

agg.cases.growth.state.plot <- ggplot(melt.use[date >= "2020-03-16"], aes(x = date, y = value, color = variable)) + 
  geom_line(aes(group = variable)) +
    labs(x = "Date",
       y = "Rate of growth (rolling 3-day window)",
       title = "Rolling 3-day rate of growth: Confirmed COVID cases",
       subtitle = use.state,
       color = "Metric") +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1), limits = 
                       c(ifelse(min(melt.use$value) > 0, 0, min(melt.use$value)), max(melt.use$value * 1.1)))
agg.cases.growth.state.plot
```

```{r message = F, warning = F}
use.state <- "Tennessee"
cases.state <- state.raw[state %in% use.state]
agg.cases.state <- cases.state[order(date),
                          .(total_cases = sum(cases, na.rm = T)),
                          by = "date"]

temp <- agg.cases.state
agg.cases.state[, total_cases_prev3 := temp$total_cases[match(agg.cases.state$date - 3, temp$date)]]
agg.cases.state[, total_cases_rtgrowth := total_cases/total_cases_prev3 - 1]
agg.cases.state[, new_cases_prev3 := total_cases - total_cases_prev3]

temp <- agg.cases.state
agg.cases.state[, new_cases_prev3_last := temp$new_cases_prev3[match(agg.cases.state$date - 1, temp$date)]]
agg.cases.state[, new_cases_rtgrowth := new_cases_prev3/new_cases_prev3_last - 1]

melt.use <- melt(agg.cases.state[, c("date", "total_cases_rtgrowth", "new_cases_rtgrowth")], id.vars = "date")
melt.use[, variable := ifelse(variable %in% "total_cases_rtgrowth", "Total Cases", "New Cases")]

agg.cases.growth.state.plot <- ggplot(melt.use[date >= "2020-03-16"], aes(x = date, y = value, color = variable)) + 
  geom_line(aes(group = variable)) +
    labs(x = "Date",
       y = "Rate of growth (rolling 3-day window)",
       title = "Rolling 3-day rate of growth: Confirmed COVID cases",
       subtitle = use.state,
       color = "Metric") +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1), limits = 
                       c(ifelse(min(melt.use$value) > 0, 0, min(melt.use$value)), max(melt.use$value * 1.1)))
agg.cases.growth.state.plot
```

### Deaths (total and new)

#### U.S.

```{r message = F, warning = F}
agg.deaths.US <- state.raw[order(date),
                          .(total_deaths = sum(deaths, na.rm = T)),
                          by = "date"]

temp <- agg.deaths.US
agg.deaths.US[, total_deaths_prev3 := temp$total_deaths[match(agg.deaths.US$date - 3, temp$date)]]
agg.deaths.US[, total_deaths_rtgrowth := total_deaths/total_deaths_prev3 - 1]
agg.deaths.US[, new_deaths_prev3 := total_deaths - total_deaths_prev3]

temp <- agg.deaths.US
agg.deaths.US[, new_deaths_prev3_last := temp$new_deaths_prev3[match(agg.deaths.US$date - 1, temp$date)]]
agg.deaths.US[, new_deaths_rtgrowth := new_deaths_prev3/new_deaths_prev3_last - 1]

melt.use <- melt(agg.deaths.US[, c("date", "total_deaths_rtgrowth", "new_deaths_rtgrowth")], id.vars = "date")
melt.use[, variable := ifelse(variable %in% "total_deaths_rtgrowth", "Total Deaths", "New Deaths")]

agg.deaths.growth.US.plot <- ggplot(melt.use[date >= "2020-03-16"], aes(x = date, y = value, color = variable)) + 
  geom_line(aes(group = variable)) +
    labs(x = "Date",
       y = "Rate of growth (rolling 3-day window)",
       title = "Rolling 3-day rate of growth: Confirmed COVID deaths",
       subtitle = "United States",
       color = "Metric") +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1), limits = 
                       c(ifelse(min(melt.use$value) > 0, 0, min(melt.use$value)), max(melt.use$value * 1.1)))
agg.deaths.growth.US.plot
```

#### Our states

```{r message = F, warning = F}
use.state <- "Washington"
deaths.state <- state.raw[state %in% use.state]
agg.deaths.state <- deaths.state[order(date),
                          .(total_deaths = sum(deaths, na.rm = T)),
                          by = "date"]

temp <- agg.deaths.state
agg.deaths.state[, total_deaths_prev3 := temp$total_deaths[match(agg.deaths.state$date - 3, temp$date)]]
agg.deaths.state[, total_deaths_rtgrowth := total_deaths/total_deaths_prev3 - 1]
agg.deaths.state[, new_deaths_prev3 := total_deaths - total_deaths_prev3]

temp <- agg.deaths.state
agg.deaths.state[, new_deaths_prev3_last := temp$new_deaths_prev3[match(agg.deaths.state$date - 1, temp$date)]]
agg.deaths.state[, new_deaths_rtgrowth := new_deaths_prev3/new_deaths_prev3_last - 1]

melt.use <- melt(agg.deaths.state[, c("date", "total_deaths_rtgrowth", "new_deaths_rtgrowth")], id.vars = "date")
melt.use[, variable := ifelse(variable %in% "total_deaths_rtgrowth", "Total Deaths", "New Deaths")]

agg.deaths.growth.state.plot <- ggplot(melt.use[date >= "2020-03-16"], aes(x = date, y = value, color = variable)) + 
  geom_line(aes(group = variable)) +
    labs(x = "Date",
       y = "Rate of growth (rolling 3-day window)",
       title = "Rolling 3-day rate of growth: Confirmed COVID deaths",
       subtitle = use.state,
       color = "Metric") +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1), limits = 
                       c(ifelse(min(melt.use$value) > 0, 0, min(melt.use$value)), max(melt.use$value * 1.1)))
agg.deaths.growth.state.plot
```

```{r message = F, warning = F}
use.state <- "California"
deaths.state <- state.raw[state %in% use.state]
agg.deaths.state <- deaths.state[order(date),
                          .(total_deaths = sum(deaths, na.rm = T)),
                          by = "date"]

temp <- agg.deaths.state
agg.deaths.state[, total_deaths_prev3 := temp$total_deaths[match(agg.deaths.state$date - 3, temp$date)]]
agg.deaths.state[, total_deaths_rtgrowth := total_deaths/total_deaths_prev3 - 1]
agg.deaths.state[, new_deaths_prev3 := total_deaths - total_deaths_prev3]

temp <- agg.deaths.state
agg.deaths.state[, new_deaths_prev3_last := temp$new_deaths_prev3[match(agg.deaths.state$date - 1, temp$date)]]
agg.deaths.state[, new_deaths_rtgrowth := new_deaths_prev3/new_deaths_prev3_last - 1]

melt.use <- melt(agg.deaths.state[, c("date", "total_deaths_rtgrowth", "new_deaths_rtgrowth")], id.vars = "date")
melt.use[, variable := ifelse(variable %in% "total_deaths_rtgrowth", "Total Deaths", "New Deaths")]

agg.deaths.growth.state.plot <- ggplot(melt.use[date >= "2020-03-16"], aes(x = date, y = value, color = variable)) + 
  geom_line(aes(group = variable)) +
    labs(x = "Date",
       y = "Rate of growth (rolling 3-day window)",
       title = "Rolling 3-day rate of growth: Confirmed COVID deaths",
       subtitle = use.state,
       color = "Metric") +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1), limits = 
                       c(ifelse(min(melt.use$value) > 0, 0, min(melt.use$value)), max(melt.use$value * 1.1)))
agg.deaths.growth.state.plot
```

```{r message = F, warning = F}
use.state <- "South Carolina"
deaths.state <- state.raw[state %in% use.state]
agg.deaths.state <- deaths.state[order(date),
                          .(total_deaths = sum(deaths, na.rm = T)),
                          by = "date"]

temp <- agg.deaths.state
agg.deaths.state[, total_deaths_prev3 := temp$total_deaths[match(agg.deaths.state$date - 3, temp$date)]]
agg.deaths.state[, total_deaths_rtgrowth := total_deaths/total_deaths_prev3 - 1]
agg.deaths.state[, new_deaths_prev3 := total_deaths - total_deaths_prev3]

temp <- agg.deaths.state
agg.deaths.state[, new_deaths_prev3_last := temp$new_deaths_prev3[match(agg.deaths.state$date - 1, temp$date)]]
agg.deaths.state[, new_deaths_rtgrowth := new_deaths_prev3/new_deaths_prev3_last - 1]
melt.use <- melt(agg.deaths.state[, c("date", "total_deaths_rtgrowth", "new_deaths_rtgrowth")], id.vars = "date")
melt.use[, variable := ifelse(variable %in% "total_deaths_rtgrowth", "Total Deaths", "New Deaths")]

agg.deaths.growth.state.plot <- ggplot(melt.use[date >= "2020-03-16"], aes(x = date, y = value, color = variable)) + 
  geom_line(aes(group = variable)) +
    labs(x = "Date",
       y = "Rate of growth (rolling 3-day window)",
       title = "Rolling 3-day rate of growth: Confirmed COVID deaths",
       subtitle = use.state,
       color = "Metric") +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1), limits = 
                       c(ifelse(min(melt.use$value) > 0, 0, min(melt.use$value)), max(melt.use$value * 1.1)))
agg.deaths.growth.state.plot
```

```{r message = F, warning = F}
use.state <- "Tennessee"
deaths.state <- state.raw[state %in% use.state]
agg.deaths.state <- deaths.state[order(date),
                          .(total_deaths = sum(deaths, na.rm = T)),
                          by = "date"]

temp <- agg.deaths.state
agg.deaths.state[, total_deaths_prev3 := temp$total_deaths[match(agg.deaths.state$date - 3, temp$date)]]
agg.deaths.state[, total_deaths_rtgrowth := total_deaths/total_deaths_prev3 - 1]
agg.deaths.state[, new_deaths_prev3 := total_deaths - total_deaths_prev3]

temp <- agg.deaths.state
agg.deaths.state[, new_deaths_prev3_last := temp$new_deaths_prev3[match(agg.deaths.state$date - 1, temp$date)]]
agg.deaths.state[, new_deaths_rtgrowth := new_deaths_prev3/new_deaths_prev3_last - 1]

melt.use <- melt(agg.deaths.state[, c("date", "total_deaths_rtgrowth", "new_deaths_rtgrowth")], id.vars = "date")
melt.use[, variable := ifelse(variable %in% "total_deaths_rtgrowth", "Total Deaths", "New Deaths")]

agg.deaths.growth.state.plot <- ggplot(melt.use[date >= "2020-03-16"], aes(x = date, y = value, color = variable)) + 
  geom_line(aes(group = variable)) +
    labs(x = "Date",
       y = "Rate of growth (rolling 3-day window)",
       title = "Rolling 3-day rate of growth: Confirmed COVID deaths",
       subtitle = use.state,
       color = "Metric") +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1), limits = 
                       c(ifelse(min(melt.use$value) > 0, 0, min(melt.use$value)), max(melt.use$value * 1.1)))
agg.deaths.growth.state.plot
```

## By population rankings

### States

#### Total confirmed cases

Table of total confirmed cases per million residents (all 50 states)
```{r message = F, warning = F}
state.per.cap <- merge(state.raw, stay_dates, by = "state", all.x = T)
state.per.cap[, casesPerMillion := as.integer(cases/population * 1000000)]
date.cut <- max(state.per.cap$date)
state.per.cap.agg <- state.per.cap[date == date.cut]
state.per.cap.agg <- state.per.cap.agg[order(-casesPerMillion), c("state", "casesPerMillion")]
state.per.cap.agg[, casesPerMillion := comma(casesPerMillion)]
state.per.cap.agg
```

```{r message = F, warning = F}
state.per.cap.agg[, row := .I]
top.ten <- state.per.cap.agg[row < 11, "state"]
state.per.cap.series <- state.per.cap[state %in% top.ten$state]
state.per.cap.plot <- ggplot(state.per.cap.series[date >= "2020-03-01"], aes(x = date, y = casesPerMillion, colour = state)) +
  geom_line(aes(group = state)) +
    labs(x = "Date",
       y = "Total confirmed COVID cases per million residents",
       title = "Total confirmed COVID cases per million residents",
       subtitle = paste0("States with 10 highest counts as of ", format(date.cut, "%B %d, %Y"))) +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) +
  scale_y_continuous(labels = comma)
state.per.cap.plot
```

#### New confirmed cases

Table of new cases per million residents: rolling 3-day total (all 50 states)
```{r message = F, warning = F}
state.per.cap <- merge(state.raw, stay_dates, by = "state", all.x = T)
date.cut <- max(state.per.cap$date)
date.min <- max(state.per.cap$date) - 3
state.per.cap.agg <- state.per.cap[date == date.cut]
temp <- state.per.cap[date == date.min]
state.per.cap.agg[, casesPrev3 := temp$cases[match(state.per.cap.agg$state, temp$state)]]
state.per.cap.agg[, newCasesPrev3 := (cases - casesPrev3)]
state.per.cap.agg[, casesNewPerMillion := as.integer(newCasesPrev3/population * 1000000)]
state.per.cap.agg <- state.per.cap.agg[order(-casesNewPerMillion), c("state", "casesNewPerMillion")]
state.per.cap.agg[, casesNewPerMillion := comma(casesNewPerMillion)]
state.per.cap.agg
```

```{r  echo = F, message = F, warning = F}
state.per.cap.agg[, row := .I]
top.ten <- state.per.cap.agg[row < 11, "state"]
state.per.cap.series <- state.per.cap[state %in% top.ten$state]
temp <- state.per.cap.series[, c("state", "date", "cases")]
setnames(temp, "cases", "casesPrev3")
state.per.cap.series[, datePrev3 := date - 3]
state.per.cap.series <- merge(state.per.cap.series, temp, all.x = T, by.x = c("state", "datePrev3"), by.y = c("state", "date"))
state.per.cap.series[, newCasesPrev3 := (cases - casesPrev3)]
state.per.cap.series[, casesNewPerMillion := as.integer(newCasesPrev3/population * 1000000)]
state.per.cap.plot <- ggplot(state.per.cap.series[date >= "2020-03-10"], aes(x = date, y = casesNewPerMillion, colour = state)) +
  geom_line(aes(group = state)) +
    labs(x = "Date",
       y = "New confirmed COVID cases per million residents",
       title = "New confirmed COVID cases per million resident (3-day rolling total)",
       subtitle = paste0("States with 10 highest counts as of ", format(date.cut, "%B %d, %Y"))) +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) +
  scale_y_continuous(labels = comma)
state.per.cap.plot
```

#### Total deaths

Table of total deaths per million residents (all 50 states)
```{r message = F, warning = F}
state.per.cap <- merge(state.raw, stay_dates, by = "state", all.x = T)
state.per.cap[, deathsPerMillion := as.integer(deaths/population * 1000000)]
date.cut <- max(state.per.cap$date)
state.per.cap.agg <- state.per.cap[date == date.cut]
state.per.cap.agg <- state.per.cap.agg[order(-deathsPerMillion), c("state", "deathsPerMillion")]
state.per.cap.agg[, deathsPerMillion := comma(deathsPerMillion)]
state.per.cap.agg
```

```{r message = F, warning = F}
state.per.cap.agg[, row := .I]
top.ten <- state.per.cap.agg[row < 11, "state"]
state.per.cap.series <- state.per.cap[state %in% top.ten$state]
state.per.cap.plot <- ggplot(state.per.cap.series[date >= "2020-03-01"], aes(x = date, y = deathsPerMillion, colour = state)) +
  geom_line(aes(group = state)) +
    labs(x = "Date",
       y = "Total COVID-related deaths per million residents",
       title = "Total COVID-related deaths per million residents",
       subtitle = paste0("States with 10 highest counts as of ", format(date.cut, "%B %d, %Y"))) +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) +
  scale_y_continuous(labels = comma)
state.per.cap.plot
```

#### New deaths

Table of new deaths per million residents: rolling 3-day total (all 50 states)
```{r message = F, warning = F}
state.per.cap <- merge(state.raw, stay_dates, by = "state", all.x = T)
date.cut <- max(state.per.cap$date)
date.min <- max(state.per.cap$date) - 3
state.per.cap.agg <- state.per.cap[date == date.cut]
temp <- state.per.cap[date == date.min]
state.per.cap.agg[, deathsPrev3 := temp$deaths[match(state.per.cap.agg$state, temp$state)]]
state.per.cap.agg[, newDeathsPrev3 := (deaths - deathsPrev3)]
state.per.cap.agg[, deathsNewPerMillion := as.integer(newDeathsPrev3/population * 1000000)]
state.per.cap.agg <- state.per.cap.agg[order(-deathsNewPerMillion), c("state", "deathsNewPerMillion")]
state.per.cap.agg[, deathsNewPerMillion := comma(deathsNewPerMillion)]
state.per.cap.agg
```

```{r message = F, warning = F}
state.per.cap.agg[, row := .I]
top.ten <- state.per.cap.agg[row < 11, "state"]
state.per.cap.series <- state.per.cap[state %in% top.ten$state]
temp <- state.per.cap.series[, c("state", "date", "deaths")]
setnames(temp, "deaths", "deathsPrev3")
state.per.cap.series[, datePrev3 := date - 3]
state.per.cap.series <- merge(state.per.cap.series, temp, all.x = T, by.x = c("state", "datePrev3"), by.y = c("state", "date"))
state.per.cap.series[, newDeathsPrev3 := (deaths - deathsPrev3)]
state.per.cap.series[, deathsNewPerMillion := as.integer(newDeathsPrev3/population * 1000000)]
state.per.cap.plot <- ggplot(state.per.cap.series[date >= "2020-03-10"], aes(x = date, y = deathsNewPerMillion, colour = state)) +
  geom_line(aes(group = state)) +
    labs(x = "Date",
       y = "New COVID-related deaths per million residents",
       title = "New COVID-related deaths per million resident (3-day rolling total)",
       subtitle = paste0("States with 10 highest counts as of ", format(date.cut, "%B %d, %Y"))) +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))
state.per.cap.plot
```

### Counties

#### Confirmed cases

Table showing total cases per million and percentile for all US counties
Includes our counties and the top 5 in the US for perspective
```{r  echo = F, message = F, warning = F}
# merge county raw and county pop
county.per.cap <- merge(county.raw, county_pop, all.x = T, by.x = c("county", "state"), by.y = c("CTYNAME", "STNAME"))
county.per.cap[, casesPerMillion := round(cases/POPESTIMATE2019 * 1000000, 2)]
date.cut <- max(county.per.cap$date)
county.per.cap <- county.per.cap[date == date.cut]
county.per.cap <- county.per.cap[order(-casesPerMillion), c("county", "state", "casesPerMillion")]
county.per.cap[, rawRanking := .I]
county.per.cap[, percentile := floor((3142 - rawRanking)/31.42)]
county.per.cap[, casesPerMillion := comma(casesPerMillion)]
county.per.cap[, flgOurCounty := ifelse(county %in% "Pierce" & state %in% "Washington", 1,
                                       ifelse(county %in% "Orange" & state %in% "California", 1,
                                              ifelse(county %in% c("Richland", "York") & state %in% "South Carolina", 1,
                                                     ifelse(county %in% "Davidson" & state %in% "Tennessee", 1, 0))))]
county.per.cap[flgOurCounty == 1 | rawRanking < 6, -c("flgOurCounty")]
```

Our county percentiles over time
```{r  echo = F, message = F, warning = F}
county.per.cap <- merge(county.raw, county_pop, all.x = T, by.x = c("county", "state"), by.y = c("CTYNAME", "STNAME"))
county.per.cap[, casesPerMillion := round(cases/POPESTIMATE2019 * 1000000, 2)]
county.per.cap <- county.per.cap[order(fips, date),
                                 .(casesPerMillion = max(casesPerMillion)),
                                 by = c("fips", "date")]

county.per.cap <- county.per.cap[!is.na(casesPerMillion) & as.Date(date) >= "2020-03-16"]
county.per.cap.casted <- dcast.data.table(county.per.cap, factor(fips) ~ factor(date), value.var = "casesPerMillion", fun.aggregate = sum)
for(i in seq.Date(as.Date("2020-03-16"), max(as.Date(county.raw$date)), by = "day")) {
 # i <- "2020-04-07" 
  i <- format(as.Date(i, origin = "1970-01-01"), "%Y-%m-%d")
    colnames <- c("fips", as.character(i))
    temp <- county.per.cap.casted[, ..colnames]
    setnames(temp, as.character(i), "casesPerMillion")
    temp <- temp[order(-casesPerMillion)]
    temp[, rawRanking := .I]
    temp[, paste0("percentile_", as.character(i)) := floor((3142 - rawRanking)/31.42)]
    county.per.cap.casted <- merge(county.per.cap.casted, temp[, c(1,4)], all.x = T, by = "fips")
}

setnames(county.per.cap.casted, "fips", "percentile_fips")
county.per.cap.casted <- county.per.cap.casted[, .SD, .SDcols = patterns("^percentile")]
county.per.cap.melt <- melt(county.per.cap.casted, id.var = "percentile_fips")
county.per.cap.melt[, variable := gsub("percentile_(.*)", "\\1", variable)]
our.fips <- c(6059, 47037, 53053, 45091, 45079)
county.per.cap.melt <- county.per.cap.melt[percentile_fips %in% our.fips]

county.per.cap.melt[, county := county.raw$county[match(county.per.cap.melt$percentile_fips, county.raw$fips)]]

plot.county.per.cap <- ggplot(county.per.cap.melt, aes(x = as.Date(variable), y = value, color = county)) +
  geom_line(aes(group = county)) +
    labs(x = "Date",
       y = "Percentile of US counties",
       title = "Our county percentiles for total COVID cases",
       subtitle = "Percentiles based on total cases per million residents",
       color = "County") +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))
plot.county.per.cap
```

#### Deaths

Table showing total deaths per million and percentile for all US counties
Includes our counties and the top 5 in the US for perspective
```{r  echo = F, message = F, warning = F}
county.per.cap <- merge(county.raw, county_pop, all.x = T, by.x = c("county", "state"), by.y = c("CTYNAME", "STNAME"))
county.per.cap[, deathsPerMillion := round(deaths/POPESTIMATE2019 * 1000000, 2)]
date.cut <- max(county.per.cap$date)
county.per.cap <- county.per.cap[date == date.cut]
county.per.cap <- county.per.cap[order(-deathsPerMillion), c("county", "state", "deathsPerMillion")]
county.per.cap[, rawRanking := .I]
county.per.cap[, percentile := floor((3142 - rawRanking)/31.42)]
county.per.cap[, deathsPerMillion := comma(deathsPerMillion)]
county.per.cap[, flgOurCounty := ifelse(county %in% "Pierce" & state %in% "Washington", 1,
                                       ifelse(county %in% "Orange" & state %in% "California", 1,
                                              ifelse(county %in% c("Richland", "York") & state %in% "South Carolina", 1,
                                                     ifelse(county %in% "Davidson" & state %in% "Tennessee", 1, 0))))]
county.per.cap[flgOurCounty == 1 | rawRanking < 6, -c("flgOurCounty")]
```

Our county percentiles over time
```{r  echo = F, message = F, warning = F}
county.per.cap <- merge(county.raw, county_pop, all.x = T, by.x = c("county", "state"), by.y = c("CTYNAME", "STNAME"))
county.per.cap[, deathsPerMillion := round(deaths/POPESTIMATE2019 * 1000000, 2)]
county.per.cap <- county.per.cap[order(fips, date),
                                 .(deathsPerMillion = max(deathsPerMillion)),
                                 by = c("fips", "date")]

county.per.cap <- county.per.cap[!is.na(deathsPerMillion) & as.Date(date) >= "2020-03-16"]
county.per.cap.casted <- dcast.data.table(county.per.cap, factor(fips) ~ factor(date), value.var = "deathsPerMillion", fun.aggregate = sum)
for(i in seq.Date(as.Date("2020-03-16"), max(as.Date(county.raw$date)), by = "day")) {
 # i <- "2020-04-07" 
  i <- format(as.Date(i, origin = "1970-01-01"), "%Y-%m-%d")
    colnames <- c("fips", as.character(i))
    temp <- county.per.cap.casted[, ..colnames]
    setnames(temp, as.character(i), "deathsPerMillion")
    temp <- temp[order(-deathsPerMillion)]
    temp[, rawRanking := .I]
    temp[, paste0("percentile_", as.character(i)) := floor((3142 - rawRanking)/31.42)]
    county.per.cap.casted <- merge(county.per.cap.casted, temp[, c(1,4)], all.x = T, by = "fips")
}

setnames(county.per.cap.casted, "fips", "percentile_fips")
county.per.cap.casted <- county.per.cap.casted[, .SD, .SDcols = patterns("^percentile")]
county.per.cap.melt <- melt(county.per.cap.casted, id.var = "percentile_fips")
county.per.cap.melt[, variable := gsub("percentile_(.*)", "\\1", variable)]
our.fips <- c(6059, 47037, 53053, 45091, 45079)
county.per.cap.melt <- county.per.cap.melt[percentile_fips %in% our.fips]

county.per.cap.melt[, county := county.raw$county[match(county.per.cap.melt$percentile_fips, county.raw$fips)]]

plot.county.per.cap <- ggplot(county.per.cap.melt, aes(x = as.Date(variable), y = value, color = county)) +
  geom_line(aes(group = county)) +
    labs(x = "Date",
       y = "Percentile of US counties",
       title = "Our county percentiles for total COVID deaths",
       subtitle = "Percentiles based on total deaths per million residents",
       color = "County") +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))
plot.county.per.cap
```

## Raw counts

### Total confirmed cases

#### U.S.

```{r message = F, warning = F}
agg.cases.US <- state.raw[order(date),
                          .(total_cases = sum(cases, na.rm = T)),
                          by = "date"]
agg.cases.US.plot <- ggplot(agg.cases.US[date >= "2020-03-01"], aes(x = date, y = total_cases)) + 
  geom_bar(stat = "identity") +
    labs(x = "Date",
       y = "Total confirmed COVID cases",
       title = "Total confirmed COVID cases",
       subtitle = "United States") +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) +
  scale_y_continuous(labels = comma)
agg.cases.US.plot
```

#### Our states

```{r message = F, warning = F}
# adjust state name to flow through rest of code (use capitalized full state name)
use.state <- "Washington"
cases.state <- state.raw[state %in% use.state]
agg.cases.state <- cases.state[order(date),
                          .(total_cases = sum(cases, na.rm = T)),
                          by = "date"]
agg.cases.state.plot <- ggplot(agg.cases.state[date >= "2020-03-01"], aes(x = date, y = total_cases)) + 
  geom_bar(stat = "identity") +
    labs(x = "Date",
       y = "Total confirmed COVID cases",
       title = "Total confirmed COVID cases",
       subtitle = use.state) +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) +
  scale_y_continuous(labels = comma)
agg.cases.state.plot
```

```{r message = F, warning = F}
# adjust state name to flow through rest of code (use capitalized full state name)
use.state <- "California"
cases.state <- state.raw[state %in% use.state]
agg.cases.state <- cases.state[order(date),
                          .(total_cases = sum(cases, na.rm = T)),
                          by = "date"]
agg.cases.state.plot <- ggplot(agg.cases.state[date >= "2020-03-01"], aes(x = date, y = total_cases)) + 
  geom_bar(stat = "identity") +
    labs(x = "Date",
       y = "Total confirmed COVID cases",
       title = "Total confirmed COVID cases",
       subtitle = use.state) +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) +
  scale_y_continuous(labels = comma)
agg.cases.state.plot
```

```{r message = F, warning = F}
# adjust state name to flow through rest of code (use capitalized full state name)
use.state <- "South Carolina"
cases.state <- state.raw[state %in% use.state]
agg.cases.state <- cases.state[order(date),
                          .(total_cases = sum(cases, na.rm = T)),
                          by = "date"]
agg.cases.state.plot <- ggplot(agg.cases.state[date >= "2020-03-01"], aes(x = date, y = total_cases)) + 
  geom_bar(stat = "identity") +
    labs(x = "Date",
       y = "Total confirmed COVID cases",
       title = "Total confirmed COVID cases",
       subtitle = use.state) +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) +
  scale_y_continuous(labels = comma)
agg.cases.state.plot
```

```{r message = F, warning = F}
# adjust state name to flow through rest of code (use capitalized full state name)
use.state <- "Tennessee"
cases.state <- state.raw[state %in% use.state]
agg.cases.state <- cases.state[order(date),
                          .(total_cases = sum(cases, na.rm = T)),
                          by = "date"]
agg.cases.state.plot <- ggplot(agg.cases.state[date >= "2020-03-01"], aes(x = date, y = total_cases)) + 
  geom_bar(stat = "identity") +
    labs(x = "Date",
       y = "Total confirmed COVID cases",
       title = "Total confirmed COVID cases",
       subtitle = use.state) +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) +
  scale_y_continuous(labels = comma)
agg.cases.state.plot
```

#### Our counties

```{r message = F, warning = F}
# adjust state and county name to flow through rest of code (use capitalized full state name)
use.state <- "Washington"
use.county <- "Pierce"
cases.county <- county.raw[state %in% use.state & county %in% use.county]
agg.cases.county <- cases.county[order(date),
                          .(total_cases = sum(cases, na.rm = T)),
                          by = "date"]
agg.cases.county.plot <- ggplot(agg.cases.county[date >= "2020-03-01"], aes(x = date, y = total_cases)) + 
  geom_bar(stat = "identity") +
    labs(x = "Date",
       y = "Total confirmed COVID cases",
       title = "Total confirmed COVID cases",
       subtitle = paste0(use.county, " County, ", use.state)) +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) +
  scale_y_continuous(labels = comma)
agg.cases.county.plot
```

```{r message = F, warning = F}
# adjust state and county name to flow through rest of code (use capitalized full state name)
use.state <- "California"
use.county <- "Orange"
cases.county <- county.raw[state %in% use.state & county %in% use.county]
agg.cases.county <- cases.county[order(date),
                          .(total_cases = sum(cases, na.rm = T)),
                          by = "date"]
agg.cases.county.plot <- ggplot(agg.cases.county[date >= "2020-03-01"], aes(x = date, y = total_cases)) + 
  geom_bar(stat = "identity") +
    labs(x = "Date",
       y = "Total confirmed COVID cases",
       title = "Total confirmed COVID cases",
       subtitle = paste0(use.county, " County, ", use.state)) +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) +
  scale_y_continuous(labels = comma)
agg.cases.county.plot
```

```{r message = F, warning = F}
# adjust state and county name to flow through rest of code (use capitalized full state name)
use.state <- "South Carolina"
use.county <- "Richland"
cases.county <- county.raw[state %in% use.state & county %in% use.county]
agg.cases.county <- cases.county[order(date),
                          .(total_cases = sum(cases, na.rm = T)),
                          by = "date"]
agg.cases.county.plot <- ggplot(agg.cases.county[date >= "2020-03-01"], aes(x = date, y = total_cases)) + 
  geom_bar(stat = "identity") +
    labs(x = "Date",
       y = "Total confirmed COVID cases",
       title = "Total confirmed COVID cases",
       subtitle = paste0(use.county, " County, ", use.state)) +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) +
  scale_y_continuous(labels = comma)
agg.cases.county.plot
```

```{r message = F, warning = F}
# adjust state and county name to flow through rest of code (use capitalized full state name)
use.state <- "South Carolina"
use.county <- "York"
cases.county <- county.raw[state %in% use.state & county %in% use.county]
agg.cases.county <- cases.county[order(date),
                          .(total_cases = sum(cases, na.rm = T)),
                          by = "date"]
agg.cases.county.plot <- ggplot(agg.cases.county[date >= "2020-03-01"], aes(x = date, y = total_cases)) + 
  geom_bar(stat = "identity") +
    labs(x = "Date",
       y = "Total confirmed COVID cases",
       title = "Total confirmed COVID cases",
       subtitle = paste0(use.county, " County, ", use.state)) +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) +
  scale_y_continuous(labels = comma)
agg.cases.county.plot
```

```{r message = F, warning = F}
# adjust state and county name to flow through rest of code (use capitalized full state name)
use.state <- "Tennessee"
use.county <- "Davidson"
cases.county <- county.raw[state %in% use.state & county %in% use.county]
agg.cases.county <- cases.county[order(date),
                          .(total_cases = sum(cases, na.rm = T)),
                          by = "date"]
agg.cases.county.plot <- ggplot(agg.cases.county[date >= "2020-03-01"], aes(x = date, y = total_cases)) + 
  geom_bar(stat = "identity") +
    labs(x = "Date",
       y = "Total confirmed COVID cases",
       title = "Total confirmed COVID cases",
       subtitle = paste0(use.county, " County, ", use.state)) +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) +
  scale_y_continuous(labels = comma)
agg.cases.county.plot
```

### New confirmed cases

#### U.S.

```{r message = F, warning = F}
agg.cases.US <- state.raw[order(date),
                          .(total_cases = sum(cases, na.rm = T)),
                          by = "date"]
temp <- agg.cases.US
agg.cases.US[, total_cases_prev := temp$total_cases[match(agg.cases.US$date - 1, temp$date)]]
agg.cases.US[, total_new_cases := total_cases - total_cases_prev]
agg.new.cases.US.plot <- ggplot(agg.cases.US[date >= "2020-03-01"], aes(x = date, y = total_new_cases)) + 
  geom_bar(stat = "identity") +
    labs(x = "Date",
       y = "Confirmed new COVID cases",
       title = "Confirmed new COVID cases by day",
       subtitle = "United States") +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) +
  scale_y_continuous(labels = comma)
agg.new.cases.US.plot
```

#### Our states

```{r message = F, warning = F}
use.state <- "Washington"
cases.state <- state.raw[state %in% use.state]
agg.cases.state <- cases.state[order(date),
                          .(total_cases = sum(cases, na.rm = T)),
                          by = "date"]
temp <- agg.cases.state
agg.cases.state[, total_cases_prev := temp$total_cases[match(agg.cases.state$date - 1, temp$date)]]
agg.cases.state[, total_new_cases := total_cases - total_cases_prev]
agg.new.cases.state.plot <- ggplot(agg.cases.state[date >= "2020-03-01"], aes(x = date, y = total_new_cases)) + 
  geom_bar(stat = "identity") +
    labs(x = "Date",
       y = "Confirmed new COVID cases",
       title = "Confirmed new COVID cases by day",
       subtitle = use.state) +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) +
  scale_y_continuous(labels = comma)
agg.new.cases.state.plot
```

```{r message = F, warning = F}
use.state <- "California"
cases.state <- state.raw[state %in% use.state]
agg.cases.state <- cases.state[order(date),
                          .(total_cases = sum(cases, na.rm = T)),
                          by = "date"]
temp <- agg.cases.state
agg.cases.state[, total_cases_prev := temp$total_cases[match(agg.cases.state$date - 1, temp$date)]]
agg.cases.state[, total_new_cases := total_cases - total_cases_prev]
agg.new.cases.state.plot <- ggplot(agg.cases.state[date >= "2020-03-01"], aes(x = date, y = total_new_cases)) + 
  geom_bar(stat = "identity") +
    labs(x = "Date",
       y = "Confirmed new COVID cases",
       title = "Confirmed new COVID cases by day",
       subtitle = use.state) +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) +
  scale_y_continuous(labels = comma)
agg.new.cases.state.plot
```

```{r message = F, warning = F}
use.state <- "South Carolina"
cases.state <- state.raw[state %in% use.state]
agg.cases.state <- cases.state[order(date),
                          .(total_cases = sum(cases, na.rm = T)),
                          by = "date"]
temp <- agg.cases.state
agg.cases.state[, total_cases_prev := temp$total_cases[match(agg.cases.state$date - 1, temp$date)]]
agg.cases.state[, total_new_cases := total_cases - total_cases_prev]
agg.new.cases.state.plot <- ggplot(agg.cases.state[date >= "2020-03-01"], aes(x = date, y = total_new_cases)) + 
  geom_bar(stat = "identity") +
    labs(x = "Date",
       y = "Confirmed new COVID cases",
       title = "Confirmed new COVID cases by day",
       subtitle = use.state) +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) +
  scale_y_continuous(labels = comma)
agg.new.cases.state.plot
```

```{r message = F, warning = F}
use.state <- "Tennessee"
cases.state <- state.raw[state %in% use.state]
agg.cases.state <- cases.state[order(date),
                          .(total_cases = sum(cases, na.rm = T)),
                          by = "date"]
temp <- agg.cases.state
agg.cases.state[, total_cases_prev := temp$total_cases[match(agg.cases.state$date - 1, temp$date)]]
agg.cases.state[, total_new_cases := total_cases - total_cases_prev]
agg.new.cases.state.plot <- ggplot(agg.cases.state[date >= "2020-03-01"], aes(x = date, y = total_new_cases)) + 
  geom_bar(stat = "identity") +
    labs(x = "Date",
       y = "Confirmed new COVID cases",
       title = "Confirmed new COVID cases by day",
       subtitle = use.state) +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) +
  scale_y_continuous(labels = comma)
agg.new.cases.state.plot
```

#### Our counties

```{r message = F, warning = F}
use.state <- "Washington"
use.county <- "Pierce"
cases.county <- county.raw[state %in% use.state & county %in% use.county]
agg.cases.county <- cases.county[order(date),
                          .(total_cases = sum(cases, na.rm = T)),
                          by = "date"]
temp <- agg.cases.county
agg.cases.county[, total_cases_prev := temp$total_cases[match(agg.cases.county$date - 1, temp$date)]]
agg.cases.county[, total_new_cases := total_cases - total_cases_prev]
agg.new.cases.county.plot <- ggplot(agg.cases.county[date >= "2020-03-01"], aes(x = date, y = total_new_cases)) + 
  geom_bar(stat = "identity") +
    labs(x = "Date",
       y = "Confirmed new COVID cases",
       title = "Confirmed new COVID cases by day",
       subtitle = paste0(use.county, " County, ", use.state)) +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) +
  scale_y_continuous(labels = comma)
agg.new.cases.county.plot
```

```{r message = F, warning = F}
use.state <- "California"
use.county <- "Orange"
cases.county <- county.raw[state %in% use.state & county %in% use.county]
agg.cases.county <- cases.county[order(date),
                          .(total_cases = sum(cases, na.rm = T)),
                          by = "date"]
temp <- agg.cases.county
agg.cases.county[, total_cases_prev := temp$total_cases[match(agg.cases.county$date - 1, temp$date)]]
agg.cases.county[, total_new_cases := total_cases - total_cases_prev]
agg.new.cases.county.plot <- ggplot(agg.cases.county[date >= "2020-03-01"], aes(x = date, y = total_new_cases)) + 
  geom_bar(stat = "identity") +
    labs(x = "Date",
       y = "Confirmed new COVID cases",
       title = "Confirmed new COVID cases by day",
       subtitle = paste0(use.county, " County, ", use.state)) +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) +
  scale_y_continuous(labels = comma)
agg.new.cases.county.plot
```

```{r message = F, warning = F}
use.state <- "South Carolina"
use.county <- "Richland"
cases.county <- county.raw[state %in% use.state & county %in% use.county]
agg.cases.county <- cases.county[order(date),
                          .(total_cases = sum(cases, na.rm = T)),
                          by = "date"]
temp <- agg.cases.county
agg.cases.county[, total_cases_prev := temp$total_cases[match(agg.cases.county$date - 1, temp$date)]]
agg.cases.county[, total_new_cases := total_cases - total_cases_prev]
agg.new.cases.county.plot <- ggplot(agg.cases.county[date >= "2020-03-01"], aes(x = date, y = total_new_cases)) + 
  geom_bar(stat = "identity") +
    labs(x = "Date",
       y = "Confirmed new COVID cases",
       title = "Confirmed new COVID cases by day",
       subtitle = paste0(use.county, " County, ", use.state)) +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) +
  scale_y_continuous(labels = comma)
agg.new.cases.county.plot
```

```{r  echo = F, message = F, warning = F}
use.state <- "South Carolina"
use.county <- "York"
cases.county <- county.raw[state %in% use.state & county %in% use.county]
agg.cases.county <- cases.county[order(date),
                          .(total_cases = sum(cases, na.rm = T)),
                          by = "date"]
temp <- agg.cases.county
agg.cases.county[, total_cases_prev := temp$total_cases[match(agg.cases.county$date - 1, temp$date)]]
agg.cases.county[, total_new_cases := total_cases - total_cases_prev]
agg.new.cases.county.plot <- ggplot(agg.cases.county[date >= "2020-03-01"], aes(x = date, y = total_new_cases)) + 
  geom_bar(stat = "identity") +
    labs(x = "Date",
       y = "Confirmed new COVID cases",
       title = "Confirmed new COVID cases by day",
       subtitle = paste0(use.county, " County, ", use.state)) +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) +
  scale_y_continuous(labels = comma)
agg.new.cases.county.plot
```

```{r message = F, warning = F}
use.state <- "Tennessee"
use.county <- "Davidson"
cases.county <- county.raw[state %in% use.state & county %in% use.county]
agg.cases.county <- cases.county[order(date),
                          .(total_cases = sum(cases, na.rm = T)),
                          by = "date"]
temp <- agg.cases.county
agg.cases.county[, total_cases_prev := temp$total_cases[match(agg.cases.county$date - 1, temp$date)]]
agg.cases.county[, total_new_cases := total_cases - total_cases_prev]
agg.new.cases.county.plot <- ggplot(agg.cases.county[date >= "2020-03-01"], aes(x = date, y = total_new_cases)) + 
  geom_bar(stat = "identity") +
    labs(x = "Date",
       y = "Confirmed new COVID cases",
       title = "Confirmed new COVID cases by day",
       subtitle = paste0(use.county, " County, ", use.state)) +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) +
  scale_y_continuous(labels = comma)
agg.new.cases.county.plot
```

### Total deaths

#### U.S.

```{r message = F, warning = F}
agg.deaths.US <- state.raw[order(date),
                          .(total_deaths = sum(deaths, na.rm = T)),
                          by = "date"]
agg.deaths.US.plot <- ggplot(agg.deaths.US[date >= "2020-03-01"], aes(x = date, y = total_deaths)) + 
  geom_bar(stat = "identity") +
    labs(x = "Date",
       y = "Total COVID-related deaths",
       title = "Total COVID-related deaths",
       subtitle = "United States") +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) +
  scale_y_continuous(labels = comma)
agg.deaths.US.plot
```

#### Our states

```{r message = F, warning = F}
# adjust state name to flow through rest of code (use capitalized full state name)
use.state <- "Washington"
deaths.state <- state.raw[state %in% use.state]
agg.deaths.state <- cases.state[order(date),
                          .(total_deaths = sum(deaths, na.rm = T)),
                          by = "date"]
agg.deaths.state.plot <- ggplot(agg.deaths.state[date >= "2020-03-01"], aes(x = date, y = total_deaths)) + 
  geom_bar(stat = "identity") +
    labs(x = "Date",
       y = "Total COVID-related deaths",
       title = "Total COVID-related deaths",
       subtitle = use.state) +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))
agg.deaths.state.plot
```

```{r message = F, warning = F}
# adjust state name to flow through rest of code (use capitalized full state name)
use.state <- "California"
deaths.state <- state.raw[state %in% use.state]
agg.deaths.state <- cases.state[order(date),
                          .(total_deaths = sum(deaths, na.rm = T)),
                          by = "date"]
agg.deaths.state.plot <- ggplot(agg.deaths.state[date >= "2020-03-01"], aes(x = date, y = total_deaths)) + 
  geom_bar(stat = "identity") +
    labs(x = "Date",
       y = "Total COVID-related deaths",
       title = "Total COVID-related deaths",
       subtitle = use.state) +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))
agg.deaths.state.plot
```

```{r message = F, warning = F}
# adjust state name to flow through rest of code (use capitalized full state name)
use.state <- "South Carolina"
deaths.state <- state.raw[state %in% use.state]
agg.deaths.state <- cases.state[order(date),
                          .(total_deaths = sum(deaths, na.rm = T)),
                          by = "date"]
agg.deaths.state.plot <- ggplot(agg.deaths.state[date >= "2020-03-01"], aes(x = date, y = total_deaths)) + 
  geom_bar(stat = "identity") +
    labs(x = "Date",
       y = "Total COVID-related deaths",
       title = "Total COVID-related deaths",
       subtitle = use.state) +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))
agg.deaths.state.plot
```

```{r message = F, warning = F}
# adjust state name to flow through rest of code (use capitalized full state name)
use.state <- "Tennessee"
deaths.state <- state.raw[state %in% use.state]
agg.deaths.state <- cases.state[order(date),
                          .(total_deaths = sum(deaths, na.rm = T)),
                          by = "date"]
agg.deaths.state.plot <- ggplot(agg.deaths.state[date >= "2020-03-01"], aes(x = date, y = total_deaths)) + 
  geom_bar(stat = "identity") +
    labs(x = "Date",
       y = "Total COVID-related deaths",
       title = "Total COVID-related deaths",
       subtitle = use.state) +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))
agg.deaths.state.plot
```

#### Our counties

```{r message = F, warning = F}
# adjust state and county name to flow through rest of code (use capitalized full state name)
use.state <- "Washington"
use.county <- "Pierce"
deaths.county <- county.raw[state %in% use.state & county %in% use.county]
agg.deaths.county <- deaths.county[order(date),
                          .(total_deaths = sum(deaths, na.rm = T)),
                          by = "date"]
agg.deaths.county.plot <- ggplot(agg.deaths.county[date >= "2020-03-01"], aes(x = date, y = total_deaths)) + 
  geom_bar(stat = "identity") +
    labs(x = "Date",
       y = "Total COVID-related deaths",
       title = "Total COVID-related deaths",
       subtitle = paste0(use.county, " County, ", use.state)) +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))
agg.deaths.county.plot
```

```{r message = F, warning = F}
# adjust state and county name to flow through rest of code (use capitalized full state name)
use.state <- "California"
use.county <- "Orange"
deaths.county <- county.raw[state %in% use.state & county %in% use.county]
agg.deaths.county <- deaths.county[order(date),
                          .(total_deaths = sum(deaths, na.rm = T)),
                          by = "date"]
agg.deaths.county.plot <- ggplot(agg.deaths.county[date >= "2020-03-01"], aes(x = date, y = total_deaths)) + 
  geom_bar(stat = "identity") +
    labs(x = "Date",
       y = "Total COVID-related deaths",
       title = "Total COVID-related deaths",
       subtitle = paste0(use.county, " County, ", use.state)) +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))
agg.deaths.county.plot
```

```{r message = F, warning = F}
# adjust state and county name to flow through rest of code (use capitalized full state name)
use.state <- "South Carolina"
use.county <- "Richland"
deaths.county <- county.raw[state %in% use.state & county %in% use.county]
agg.deaths.county <- deaths.county[order(date),
                          .(total_deaths = sum(deaths, na.rm = T)),
                          by = "date"]
agg.deaths.county.plot <- ggplot(agg.deaths.county[date >= "2020-03-01"], aes(x = date, y = total_deaths)) + 
  geom_bar(stat = "identity") +
    labs(x = "Date",
       y = "Total COVID-related deaths",
       title = "Total COVID-related deaths",
       subtitle = paste0(use.county, " County, ", use.state)) +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))
agg.deaths.county.plot
```

```{r message = F, warning = F}
# adjust state and county name to flow through rest of code (use capitalized full state name)
use.state <- "South Carolina"
use.county <- "York"
deaths.county <- county.raw[state %in% use.state & county %in% use.county]
agg.deaths.county <- deaths.county[order(date),
                          .(total_deaths = sum(deaths, na.rm = T)),
                          by = "date"]
agg.deaths.county.plot <- ggplot(agg.deaths.county[date >= "2020-03-01"], aes(x = date, y = total_deaths)) + 
  geom_bar(stat = "identity") +
    labs(x = "Date",
       y = "Total COVID-related deaths",
       title = "Total COVID-related deaths",
       subtitle = paste0(use.county, " County, ", use.state)) +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))
agg.deaths.county.plot
```

```{r message = F, warning = F}
# adjust state and county name to flow through rest of code (use capitalized full state name)
use.state <- "Tennessee"
use.county <- "Davidson"
deaths.county <- county.raw[state %in% use.state & county %in% use.county]
agg.deaths.county <- deaths.county[order(date),
                          .(total_deaths = sum(deaths, na.rm = T)),
                          by = "date"]
agg.deaths.county.plot <- ggplot(agg.deaths.county[date >= "2020-03-01"], aes(x = date, y = total_deaths)) + 
  geom_bar(stat = "identity") +
    labs(x = "Date",
       y = "Total COVID-related deaths",
       title = "Total COVID-related deaths",
       subtitle = paste0(use.county, " County, ", use.state)) +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))
agg.deaths.county.plot
```

### New deaths

#### U.S.

```{r message = F, warning = F}
agg.deaths.US <- state.raw[order(date),
                          .(total_deaths = sum(deaths, na.rm = T)),
                          by = "date"]
temp <- agg.deaths.US
agg.deaths.US[, total_deaths_prev := temp$total_deaths[match(agg.deaths.US$date - 1, temp$date)]]
agg.deaths.US[, total_new_deaths := total_deaths - total_deaths_prev]
agg.new.deaths.US.plot <- ggplot(agg.deaths.US[date >= "2020-03-01"], aes(x = date, y = total_new_deaths)) + 
  geom_bar(stat = "identity") +
    labs(x = "Date",
       y = "New COVID-related deaths",
       title = "New COVID-related deaths by day",
       subtitle = "United States") +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) +
  scale_y_continuous(labels = comma)
agg.new.deaths.US.plot
```

#### Our states

```{r message = F, warning = F}
use.state <- "Washington"
deaths.state <- state.raw[state %in% use.state]
agg.deaths.state <- deaths.state[order(date),
                          .(total_deaths = sum(deaths, na.rm = T)),
                          by = "date"]
temp <- agg.deaths.state
agg.deaths.state[, total_deaths_prev := temp$total_deaths[match(agg.deaths.state$date - 1, temp$date)]]
agg.deaths.state[, total_new_deaths := total_deaths - total_deaths_prev]
agg.new.deaths.state.plot <- ggplot(agg.deaths.state[date >= "2020-03-01"], aes(x = date, y = total_new_deaths)) + 
  geom_bar(stat = "identity") +
    labs(x = "Date",
       y = "New COVID-related deaths",
       title = "New COVID-related deaths by day",
       subtitle = use.state) +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))
agg.new.deaths.state.plot
```

```{r message = F, warning = F}
use.state <- "California"
deaths.state <- state.raw[state %in% use.state]
agg.deaths.state <- deaths.state[order(date),
                          .(total_deaths = sum(deaths, na.rm = T)),
                          by = "date"]
temp <- agg.deaths.state
agg.deaths.state[, total_deaths_prev := temp$total_deaths[match(agg.deaths.state$date - 1, temp$date)]]
agg.deaths.state[, total_new_deaths := total_deaths - total_deaths_prev]
agg.new.deaths.state.plot <- ggplot(agg.deaths.state[date >= "2020-03-01"], aes(x = date, y = total_new_deaths)) + 
  geom_bar(stat = "identity") +
    labs(x = "Date",
       y = "New COVID-related deaths",
       title = "New COVID-related deaths by day",
       subtitle = use.state) +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))
agg.new.deaths.state.plot
```

```{r message = F, warning = F}
use.state <- "South Carolina"
deaths.state <- state.raw[state %in% use.state]
agg.deaths.state <- deaths.state[order(date),
                          .(total_deaths = sum(deaths, na.rm = T)),
                          by = "date"]
temp <- agg.deaths.state
agg.deaths.state[, total_deaths_prev := temp$total_deaths[match(agg.deaths.state$date - 1, temp$date)]]
agg.deaths.state[, total_new_deaths := total_deaths - total_deaths_prev]
agg.new.deaths.state.plot <- ggplot(agg.deaths.state[date >= "2020-03-01"], aes(x = date, y = total_new_deaths)) + 
  geom_bar(stat = "identity") +
    labs(x = "Date",
       y = "New COVID-related deaths",
       title = "New COVID-related deaths by day",
       subtitle = use.state) +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))
agg.new.deaths.state.plot
```

```{r message = F, warning = F}
use.state <- "Tennessee"
deaths.state <- state.raw[state %in% use.state]
agg.deaths.state <- deaths.state[order(date),
                          .(total_deaths = sum(deaths, na.rm = T)),
                          by = "date"]
temp <- agg.deaths.state
agg.deaths.state[, total_deaths_prev := temp$total_deaths[match(agg.deaths.state$date - 1, temp$date)]]
agg.deaths.state[, total_new_deaths := total_deaths - total_deaths_prev]
agg.new.deaths.state.plot <- ggplot(agg.deaths.state[date >= "2020-03-01"], aes(x = date, y = total_new_deaths)) + 
  geom_bar(stat = "identity") +
    labs(x = "Date",
       y = "New COVID-related deaths",
       title = "New COVID-related deaths by day",
       subtitle = use.state) +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))
agg.new.deaths.state.plot
```

#### Our counties

```{r message = F, warning = F}
use.state <- "Washington"
use.county <- "Pierce"
deaths.county <- county.raw[state %in% use.state & county %in% use.county]
agg.deaths.county <- deaths.county[order(date),
                          .(total_deaths = sum(deaths, na.rm = T)),
                          by = "date"]
temp <- agg.deaths.county
agg.deaths.county[, total_deaths_prev := temp$total_deaths[match(agg.deaths.county$date - 1, temp$date)]]
agg.deaths.county[, total_new_deaths := total_deaths - total_deaths_prev]
agg.new.deaths.county.plot <- ggplot(agg.deaths.county[date >= "2020-03-01"], aes(x = date, y = total_new_deaths)) + 
  geom_bar(stat = "identity") +
    labs(x = "Date",
       y = "New COVID-related deaths",
       title = "New COVID-related deaths by day",
       subtitle = paste0(use.county, " County, ", use.state)) +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))
agg.new.deaths.county.plot
```

```{r message = F, warning = F}
use.state <- "California"
use.county <- "Orange"
deaths.county <- county.raw[state %in% use.state & county %in% use.county]
agg.deaths.county <- deaths.county[order(date),
                          .(total_deaths = sum(deaths, na.rm = T)),
                          by = "date"]
temp <- agg.deaths.county
agg.deaths.county[, total_deaths_prev := temp$total_deaths[match(agg.deaths.county$date - 1, temp$date)]]
agg.deaths.county[, total_new_deaths := total_deaths - total_deaths_prev]
agg.new.deaths.county.plot <- ggplot(agg.deaths.county[date >= "2020-03-01"], aes(x = date, y = total_new_deaths)) + 
  geom_bar(stat = "identity") +
    labs(x = "Date",
       y = "New COVID-related deaths",
       title = "New COVID-related deaths by day",
       subtitle = paste0(use.county, " County, ", use.state)) +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))
agg.new.deaths.county.plot
```

```{r message = F, warning = F}
use.state <- "South Carolina"
use.county <- "Richland"
deaths.county <- county.raw[state %in% use.state & county %in% use.county]
agg.deaths.county <- deaths.county[order(date),
                          .(total_deaths = sum(deaths, na.rm = T)),
                          by = "date"]
temp <- agg.deaths.county
agg.deaths.county[, total_deaths_prev := temp$total_deaths[match(agg.deaths.county$date - 1, temp$date)]]
agg.deaths.county[, total_new_deaths := total_deaths - total_deaths_prev]
agg.new.deaths.county.plot <- ggplot(agg.deaths.county[date >= "2020-03-01"], aes(x = date, y = total_new_deaths)) + 
  geom_bar(stat = "identity") +
    labs(x = "Date",
       y = "New COVID-related deaths",
       title = "New COVID-related deaths by day",
       subtitle = paste0(use.county, " County, ", use.state)) +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))
agg.new.deaths.county.plot
```

```{r message = F, warning = F}
use.state <- "South Carolina"
use.county <- "York"
deaths.county <- county.raw[state %in% use.state & county %in% use.county]
agg.deaths.county <- deaths.county[order(date),
                          .(total_deaths = sum(deaths, na.rm = T)),
                          by = "date"]
temp <- agg.deaths.county
agg.deaths.county[, total_deaths_prev := temp$total_deaths[match(agg.deaths.county$date - 1, temp$date)]]
agg.deaths.county[, total_new_deaths := total_deaths - total_deaths_prev]
agg.new.deaths.county.plot <- ggplot(agg.deaths.county[date >= "2020-03-01"], aes(x = date, y = total_new_deaths)) + 
  geom_bar(stat = "identity") +
    labs(x = "Date",
       y = "New COVID-related deaths",
       title = "New COVID-related deaths by day",
       subtitle = paste0(use.county, " County, ", use.state)) +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))
agg.new.deaths.county.plot
```

```{r message = F, warning = F}
use.state <- "Tennessee"
use.county <- "Davidson"
deaths.county <- county.raw[state %in% use.state & county %in% use.county]
agg.deaths.county <- deaths.county[order(date),
                          .(total_deaths = sum(deaths, na.rm = T)),
                          by = "date"]
temp <- agg.deaths.county
agg.deaths.county[, total_deaths_prev := temp$total_deaths[match(agg.deaths.county$date - 1, temp$date)]]
agg.deaths.county[, total_new_deaths := total_deaths - total_deaths_prev]
agg.new.deaths.county.plot <- ggplot(agg.deaths.county[date >= "2020-03-01"], aes(x = date, y = total_new_deaths)) + 
  geom_bar(stat = "identity") +
    labs(x = "Date",
       y = "New COVID-related deaths",
       title = "New COVID-related deaths by day",
       subtitle = paste0(use.county, " County, ", use.state)) +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))
agg.new.deaths.county.plot
```

## Stay-at-home comparisons
* The section is a work in progress -- I'll probably change and add plots over the coming days.
* I'm trying to see what effect stay-at-home orders have on spread. It's inherentely messy because some states have localized orders but not orders for their full state and because states implemented the orders at different times.
* The current plot shows the rolling 3-day average of new cases per million residents, averaged across 2 groups of states -- those with a state-wide stay-at-home order in effect and those without. The size of the dots reflect the number of states in each group for each given day (so the dots are getting bigger for the group with orders over time, as more states implement those orders).
* So far, this view doesn't show if these orders are "working" or not. Instead, the biggest takeaway from the current plot is that the states that implemented orders are reacting to having worse spread, which is why the case rate is so much higher at the start for them. As a state decides it has a problem and implements an order, it joins the grouping of states with orders and drives the rate up for that group.
* Over time (if the policies work), what we'd expect is that the line for the group with orders flattens a bit, but we'll see. It also may become moot if all states issue orders.
* The next plot I want to add will show rate of spread based on how long an order has been in effect, but still thinking about how to do this. Mostly, I'm trying to decide what the right baseline is to compare against.

```{r message = F, warning = F}
state.dt <- state.raw
state.dt[, stay_date := stay_dates$start_date[match(state.raw$state, stay_dates$state)]]
state.dt[, population := stay_dates$population[match(state.raw$state, stay_dates$state)]]
#state.dt[, flgOrder := ifelse(is.na(stay_date), 0, 1)]
state.dt[, flgOrder := ifelse(is.na(stay_date) | stay_date >= date, "No", "Yes")]
#state.dt[, ctDaysFromBaseline := ifelse(flgOrder == 0, as.integer(date - as.Date("2020-03-10")),
#                                         as.integer(date - stay_date))]
#state.dt[, ctDaysFromBaseline := as.integer(date - stay_date)]
temp <- state.dt
temp <- temp[, c("date", "state", "cases")]
setnames(temp, "cases", "cases_prev3")
state.dt[, dtPrev3 := date - 3]
state.dt <- merge(state.raw, temp, all.x = T, by.x = c("state", "dtPrev3"), by.y = c("state", "date"))
state.dt[, casesNewPerMillion := as.integer((cases - cases_prev3)/population * 1000000)]
state.agg.comp <- state.dt[date >= "2020-03-10" & !state %in% "Puerto Rico"]
state.agg.comp <- state.agg.comp[order(date, flgOrder),
                          .(avgNewCases = mean(casesNewPerMillion, na.rm = T),
                            cts = .N),
                           by = c("date", "flgOrder")]
state.agg.comp.plot <- ggplot(state.agg.comp, aes(x = date, y = avgNewCases, colour = as.factor(flgOrder))) +
  geom_point(aes(size = cts)) +
  geom_line(aes(group = flgOrder)) +
    labs(x = "Date",
       y = "New confirmed COVID cases per million residents",
       title = "New COVID cases vs.stay-at-home order",
       subtitle = "Average of 3-day rolling totals per million residents",
       size = "Count of states",
       colour = "Stay-at-home order") +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))
state.agg.comp.plot
```
